---
title: "06_HCA_sample_QC"
author: "Luise A. Seeker"
date: "05/02/2021"
output: html_document
---

# Sample quality control

The purpose of this script is to test sample quality and flag those samples
for removal that we do not wish to keep for downstream analyses.



```{r}
library(Seurat)
library(ggplot2)
library(ggsci)
library(scales)
library(RCurl)
library(AnnotationHub)
library(dplyr)

```

Pick colour paletts

```{r}

mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)

mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)

show_col(mycoloursP, labels =F)
```

```{r}

seur_comb <- readRDS("/Users/lseeker/Documents/Work/HumanCellAtlas/splice_control_out/datasets/04_scran_normalised/combined_SCE/combined_seur_norm_raw.RDS")

```


```{r}

# this is plotting percent of mito genes that was etimated before spliced and 
# unspliced was merged
Idents(seur_comb) <- "Tissue"

VlnPlot(seur_comb, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 3, pt.size = 0.1, cols = mycoloursP[3:40])


VlnPlot(seur_comb, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 3, pt.size = 0, cols = mycoloursP[3:40])
```



```{r, fig.width=5, fig.height=3, fig.fullwidth=TRUE}

plot1 <- FeatureScatter(seur_comb, feature1 = "nCount_RNA", 
                        feature2 = "percent.mt", cols = mycoloursP[3:40])
plot2 <- FeatureScatter(seur_comb, feature1 = "nCount_RNA", 
                        feature2 = "nFeature_RNA", cols = mycoloursP[3:40])
plot1 + plot2

```


The mitochondrial percentage was based on spliced counts only. After adding the
unspliced that don't contain mitochondrial genes this percentage should go down. 
For some reason the standard Seurat way of doing this gives wrong results. 
So here I do it manually:

```{r}

df <- data.frame(genes = rownames(seur_comb))
df[grep("^MT-", df$genes),]




mt_genes <- rownames(seur_comb)[grep("^MT-",rownames(seur_comb))]
C <- GetAssayData(object = seur_comb, slot = "counts")

percent_mito <- colSums(C[mt_genes,])/Matrix::colSums(C)*100
seur_comb <- AddMetaData(seur_comb, percent_mito, col.name = "total_percent_mito")



rb_genes <- rownames(seur_comb)[grep("^RP[SL]",rownames(seur_comb))]
percent_ribo <- colSums(C[rb_genes,])/Matrix::colSums(C)*100
seur_comb <- AddMetaData(seur_comb, percent_ribo, col.name = "percent_ribo")



VlnPlot(seur_comb, features = "total_percent_mito", pt.size = 0, cols = 
          mycoloursP) + NoLegend()
```


```{r}

VlnPlot(seur_comb, features = "percent_ribo", pt.size = 0, cols = mycoloursP) + 
  NoLegend()



test <- PercentageFeatureSet(object = seur_comb, pattern = "^MT-")
```

I can't find convincing literature on filtering based on ribosomal genes. So
I do not filter here.

```{r}
Idents(seur_comb) <- "Tissue"
plot1 <- FeatureScatter(seur_comb, feature1 = "nCount_RNA", 
                        feature2 = "total_percent_mito", 
                        cols = mycoloursP[3:40])
plot2 <- FeatureScatter(seur_comb, feature1 = "nCount_RNA", 
                        feature2 = "nFeature_RNA", cols = mycoloursP[3:40])
plot1 + plot2

```

Some more quality checks:
```{r}
summaryStats <- as.data.frame(tapply(seur_comb$nCount_RNA, 
                                   seur_comb@meta.data$ProcessNumber, mean))
names(summaryStats) <- "MeanUMI"
summaryStats$ProcessNumber <- rownames(summaryStats)
rownames(summaryStats) <- NULL

medianUMI<-as.data.frame(tapply(seur_comb$nCount_RNA, 
                                seur_comb@meta.data$ProcessNumber, median))
names(medianUMI) <- "MedianUMI"

summaryStats$MedianUMI <- medianUMI$MedianUMI


meanFeatures <- as.data.frame(tapply(seur_comb$nFeature_RNA, 
                                     seur_comb@meta.data$ProcessNumber, mean))
names(meanFeatures) <- "meanFeatures"

summaryStats$meanFeatures <- meanFeatures$meanFeatures


medianFeatures <- as.data.frame(tapply(seur_comb$nFeature_RNA, 
                                       seur_comb@meta.data$ProcessNumber, 
                                       median))

names(medianFeatures) <- "medianFeatures"

summaryStats$medianFeatures <- medianFeatures$medianFeatures


maxUMIs <- as.data.frame(tapply(seur_comb$nCount_RNA, 
                                seur_comb@meta.data$ProcessNumber, max))
names(maxUMIs)<-"maxUMI"


minUMI <- as.data.frame(tapply(seur_comb$nCount_RNA, 
                               seur_comb@meta.data$ProcessNumber, min))
names(minUMI) <- "minUMI"

maxFeatures <- as.data.frame(tapply(seur_comb$nFeature_RNA, 
                                    seur_comb@meta.data$ProcessNumber, max))
names(maxFeatures) <- "maxFeatures"

minFeatures <- as.data.frame(tapply(seur_comb$nFeature_RNA, 
                                    seur_comb@meta.data$ProcessNumber, 
                                    min))
names(minFeatures)<-"minFeatures"


summaryStats$maxUMIs<-maxUMIs$maxUMI
summaryStats$minUMIs<-minUMI$minUMI
summaryStats$minFeatures<-minFeatures$minFeatures
summaryStats$maxFeatures<-maxFeatures$maxFeatures


hist(summaryStats$MeanUMI, col = mycoloursP[3], breaks = 10)
hist(summaryStats$MedianUMI, col = mycoloursP[4], breaks = 10)
hist(summaryStats$medianFeatures, col = mycoloursP[5], breaks = 10)
hist(summaryStats$meanFeatures, col = mycoloursP[6],breaks = 10)
hist(summaryStats$minFeatures, col = mycoloursP[7],breaks = 10)
hist(summaryStats$maxFeatures, col = mycoloursP[8],breaks = 10)
hist(summaryStats$minUMIs, col = mycoloursP[9],breaks = 10)
hist(summaryStats$maxUMIs, col = mycoloursP[10],breaks = 10)

```

# added later when more filtering criteria were known:

```{r, eval = F}
qc_dat <- data.frame(ProcessNumber = seur_comb@meta.data$process_number,
                    QC = seur_comb@meta.data$sample_qc_failed)
qc_dat <- qc_dat[!(duplicated(qc_dat$ProcessNumber)),]

summary_stats <- merge(summaryStats, qc_dat, by = "ProcessNumber")
summary_stats$QC_all <- ifelse(summary_stats$QC == "TRUE", "exclude", "include")


rin <- data.frame(ProcessNumber = seur_comb@meta.data$process_number,
                    RIN = seur_comb@meta.data$RIN)

rin <- rin[!(duplicated(rin$ProcessNumber)),]



summary_stats <- merge(summary_stats, rin, by = "ProcessNumber")

summary_stats$RIN <- as.numeric(summary_stats$RIN)



cor<-cor.test(summary_stats$RIN, summary_stats$MeanUMI)
r<-round(cor$estimate, 2)
p<-round(cor$p.value, 3)
p<-ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))


ggplot(summary_stats, aes(x=MeanUMI, y= RIN, colour= QC_all))+ 
  geom_point(size = 3) + xlab("MeanUMI") + 
  ylab ("RIN") + theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) 

# not all samples have RIN values, RNA isolation kits were impossible to obtain 
# during the COVID19 pandemic
```




```{r}



summaryStats$QC_UMI<-ifelse(summaryStats$MeanUMI < 500, "exclude", 
                            "include")

cor<-cor.test(summaryStats$MeanUMI, summaryStats$MedianUMI)
r<-round(cor$estimate, 2)
p<-round(cor$p.value, 3)
p<-ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))


ggplot(summaryStats, aes(x=MeanUMI, y= MedianUMI, colour= QC_UMI))+ 
  geom_point(size = 3) + xlab("Mean UMI") + 
  ylab ("Median UMI") + theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) +
  geom_smooth(method = "lm", se = T) + 
  annotate("text", x = 650, y = 1800, label = 
             paste("r = ", r, "  p ", p, sep=""),
           size = 8)+
  xlim(0, max(summaryStats$MeanUMI) + 50)


ggplot(summaryStats, aes(x=MeanUMI, y= MedianUMI, colour= QC_UMI))+ 
  geom_point(size = 3) + xlab("Mean UMI") + 
  ylab ("Median UMI") + theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) +
  geom_smooth(method = "lm", se = T) + 
  annotate("text", x = 850, y = 1800, label = 
             paste("r = ", r, "  p ", p, sep=""),
           size = 7)+
  xlim(0, max(summaryStats$MeanUMI) + 50) +
  theme(legend.position = "none")
  




```





  
```{r}

cor <- cor.test(summaryStats$meanFeatures, summaryStats$medianFeatures)
r <- round(cor$estimate, 2)
p <- round(cor$p.value, 3)
p <- ifelse(p == 0, p <- "< 0.001", p <- paste( " = " , p, sep=""))

ggplot(summaryStats, aes(x=meanFeatures, 
                         y= medianFeatures, 
                         colour= QC_UMI))+geom_point(size = 3) + 
  xlab("Mean features") + 
  ylab ("Median features") + 
  theme_classic() + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  geom_smooth(method = "lm", se = T) + 
  annotate("text", x = 400, y = 600, label = 
             paste("r = ", r, "  p ", p, sep=""))


ggplot(summaryStats, aes(x=meanFeatures, 
                         y= medianFeatures, 
                         colour= QC_UMI))+geom_point(size = 3) + 
  xlab("Mean features") + 
  ylab ("Median features") + 
  theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  geom_smooth(method = "lm", se = T) + 
  annotate("text", x = 700, y = 1200, label = 
             paste("r = ", r, "  p ", p, sep=""),
           size = 7) +
  theme(legend.position = "none")
  






```


```{r}
cor <- cor.test(summaryStats$MedianUMI, summaryStats$medianFeatures)
r <- round(cor$estimate, 2)
p <- round(cor$p.value, 3)
p <- ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))

ggplot(summaryStats, aes(x=MedianUMI, y= medianFeatures, colour= QC_UMI)) +
  geom_point(size = 3)+ xlab("Median UMI") + 
  ylab ("Median features") + 
  theme_classic() + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  geom_smooth(method = "lm", se = T)+ 
  annotate("text", x = 400, y = 600, label = paste("r = ", r, "  p ", p, sep=""))

ggplot(summaryStats, aes(x=MedianUMI, y= medianFeatures, colour= QC_UMI)) +
  geom_point(size = 3)+ xlab("Median UMI") + 
  ylab ("Median features") + 
  theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  geom_smooth(method = "lm", se = T)+ 
  annotate("text", x = 800, y = 1250, label = paste("r = ", r, "  p ", p, sep=""),
           size = 7)+
  theme(legend.position = "none")

  

```


```{r}

cor<-cor.test(summaryStats$MeanUMI, summaryStats$meanFeatures)
r<-round(cor$estimate, 2)
p<-round(cor$p.value, 3)
p<-ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))


ggplot(summaryStats, aes(x=MeanUMI, 
                         y= meanFeatures, 
                         colour= QC_UMI)) + 
  geom_point(size = 3) + 
  xlab("Mean UMI") + 
  ylab ("Mean features") + 
  theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  geom_smooth(method = "lm", se = T) + 
  annotate("text", x = 600, y = 850, 
           label = paste("r = ", r, "  p ", p, sep="",size = 7)) +
  theme(legend.position = "none")


ggplot(summaryStats, aes(x=MeanUMI, 
                         y= meanFeatures, 
                         colour= QC_UMI)) + 
  geom_point(size = 3) + 
  xlab("Mean UMI") + 
  ylab ("Mean features") + 
  theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  geom_smooth(method = "lm", se = T) + 
  annotate("text", x = 1100, y = 1200, 
           label = paste("r = ", r, "  p ", p, sep=""),
           size = 7) +
  theme(legend.position = "none")


```


```{r}

cor<-cor.test(summaryStats$minUMIs, summaryStats$maxUMIs)
r<-round(cor$estimate, 2)
p<-round(cor$p.value, 3)
p<-ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))

ggplot(summaryStats, aes(x=minUMIs, 
                         y= maxUMIs, 
                         colour= QC_UMI)) + 
  geom_point(size = 3)+ xlab("Min UMI") + 
  ylab ("Max UMI") + 
  theme_classic() + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  annotate("text", x = 250, y = 6000, 
           label = paste("r = ", r, "  p ", p, sep=""))+ 
  geom_smooth(method = "lm", se = F)


```


```{r}

cor<-cor.test(summaryStats$minFeatures, summaryStats$maxFeatures)
r<-round(cor$estimate, 2)
p<-round(cor$p.value, 3)
p<-ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))

ggplot(summaryStats, aes(x=minFeatures, 
                         y= maxFeatures, 
                         colour= QC_UMI)) + 
  geom_point(size = 3) + 
  xlab("Min Features") + 
  ylab ("Max features") + 
  theme_classic() + 
  scale_color_manual(values = mycoloursP[16:17]) + 
  annotate("text", x = 100, y = 4200, 
           label = paste("r = ", r, "  p ", p, sep=""))


```



```{r}
dir.create("/Users/lseeker/Documents/Work/HumanCellAtlas/splice_control_out/summary_stats_after_qc")
write.csv(summaryStats,"/Users/lseeker/Documents/Work/HumanCellAtlas/splice_control_out/summary_stats_after_qc/sum_stats_after_qc.csv")
          


```


```{r}
sum_stats <- merge(seur_comb@meta.data, summaryStats, by = "ProcessNumber" )
barcode_oder <- seur_comb@meta.data$Barcode
sum_stats <- sum_stats[match(barcode_oder, sum_stats$Barcode),]
head(sum_stats$Barcode == seur_comb@meta.data$Barcode)

qc_umi_vector <- sum_stats$QC_UMI

seur_comb@meta.data$QC_UMI <- qc_umi_vector


extr_RIN <- sum_stats[!duplicated(sum_stats$process_number),]
```

```{r}
extr_RIN <- subset(extr_RIN, extr_RIN$RINvalue != "NULL")

cor<-cor.test(extr_RIN$MeanUMI, as.numeric(paste(extr_RIN$RIN)))
r<-round(cor$estimate, 2)
p<-round(cor$p.value, 3)
p<-ifelse(p==0, p<- "< 0.001", p<-paste( " = " , p, sep=""))

ggplot(extr_RIN, aes(x=MeanUMI, 
                         y= RINvalue, 
                         colour= QC_UMI)) + 
  geom_point(size = 3)+ xlab("Mean UMI") + 
  ylab ("RIN") + 
  theme_classic(20) + 
  scale_color_manual(values= mycoloursP[16:17]) + 
  annotate("text", x = 1100, y = 27, 
           label = paste("r = ", r, "  p ", p, sep=""),
           size = 7) +
  theme(legend.position = "none")



```



```{r}


met_data<- seur_comb@meta.data
d <- met_data %>% 
  group_by(process_number) %>%
  summarise(no_rows = length(process_number))

met_data <- merge(met_data, d, by = "process_number")

u_met_dat <- met_data[!duplicated(met_data$process_number),]


u_met_dat <- subset(u_met_dat, u_met_dat$RIN != "NULL")

ggplot(u_met_dat, aes(x = no_rows, y = RIN, 
                      colour = QC_UMO)) + 
  geom_point() + 
  theme_classic(20) + 
  xlab("Nuclei count") + 
  scale_color_manual(values= mycoloursP[16:17])+
  theme(legend.position = "none")

```



```{r}

ggplot(u_met_dat, aes(x = no_rows, y = RIN, 
                      colour = sample_qc_failed)) + 
  geom_point() + 
  theme_classic(20) + 
  xlab("Nuclei count")+ 
  scale_color_manual(values= c(mycoloursP[6], mycoloursP[5])) +
  theme(legend.position = "none")


```


```{r}

sessionInfo()

```