---
title: "08_CellCycle_annotation"
author: "Luise A. Seeker"
date: "05/02/2021"
output: html_document
---
For the cluster quality control it may be helpful to identify mitotic cells. 
Here the standard Seurat method for cell cycle annotation is used. 

https://satijalab.org/seurat/archive/v3.1/cell_cycle_vignette.html

```{r}
library(Seurat)
library(ggplot2)
library(ggsci)
library("scales")
library(RCurl)
library(AnnotationHub)


```

Pick colour paletts

```{r}

mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)

mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)

show_col(mycoloursP, labels =F)
```


```{r}

seur_comb <- readRDS("/Users/lseeker/Documents/Work/HumanCellAtlas/splice_control_out/datasets/04_scran_normalised/combined_SCE/combined_seur_norm_raw.RDS")

```


Annotate for cell cycle

```{r}
exp_mat <- read.table(file = "/Users/lseeker/Documents/Work/HumanCellAtlas/CellCycleAnnotationFiles/nestorawa_forcellcycle_expressionMatrix.txt", header = TRUE, 
    as.is = TRUE, row.names = 1)

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s_genes <- cc.genes$s.genes
g2m_genes <- cc.genes$g2m.genes


```

```{r}

seur_comb <- CellCycleScoring(seur_comb, s.features = s_genes, 
                              g2m.features = g2m_genes, 
                              set.ident = TRUE)

# view cell cycle scores and phase assignments
head(seur_comb$Phase)


```

```{r, fig.width=7, fig.height=6, fig.fullwidth=TRUE}
DimPlot(seur_comb, 
        label = F, 
        cols = mycoloursP[2:50], 
        group.by = "Phase")


```

```{r, fig.width=10, fig.height=6, fig.fullwidth=TRUE}
DimPlot(seur_comb, 
        label = F, 
        cols = mycoloursP[2:50], 
        group.by = "Phase",
        split.by = "Phase")


```

Plot S-phase genes
```{r, fig.width=10, fig.height=50, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = s_genes,
            ncol = 2)



```


Plot G2M-phase genes
```{r, fig.width=10, fig.height=60, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = g2m_genes,
            ncol = 2)



```


I think the annotation does not work well for our CNS samples. It indicates
that a lot more cells are activel dividing than there should be. 
And looking at the genes used for the cell cycle annotation, a lot of them are
not expressed in our tissue.


```{r}

sessionInfo()
```