---
title: "09_Marker_gene_expr_test"
author: "Luise A. Seeker"
date: "05/02/2021"
output: html_document
---

```{r}
library(Seurat)
library(ggplot2)
library(ggsci)
library("scales")
library(RCurl)
library(AnnotationHub)


```

Pick colour paletts

```{r}

mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)

mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)

show_col(mycoloursP, labels =F)
```


```{r}

seur_comb <- readRDS("/Users/lseeker/Documents/Work/HumanCellAtlas/splice_control_out/datasets/04_scran_normalised/combined_SCE/combined_seur_norm_raw.RDS")

```


Neurons:
```{r, fig.width=7, fig.height=5, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("SNAP25", "STMN2", "RBFOX3", "GABRB2"))
```

Inhibitory neurons
```{r, fig.width=7, fig.height=5, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("GAD1", "GAD2", "SLC32A1", "PVALB"))



```
Add inhibitory neurons

```{r, fig.width=7, fig.height=5, fig.fullwidth=TRUE}

FeaturePlot(seur_comb, features = c("ADARB2", "VIP", "LAMP5", "LHX6"))


```
Add 2 Inhibitory Neurons

```{r, fig.width=7, fig.height=23, fig.fullwidth=TRUE}

FeaturePlot(seur_comb, features = c("LAMA3", "VIP", "TAC3", "RYR3", "TSHZ2",
                                    "RELN", "IL1RAPL2", "CXCL14", "EYA4",
                                    "FBXL7", "KIT", "MEIS2", "PBX3", "PLCL1", 
                                    "MYO5B", "TRHDE", "PLCH1"), 
            ncol = 2)


```


Excitatory neurons
```{r, fig.width=7, fig.height=5, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("SATB2", "SLC12A6", "SLC17A7"))

```
Add exitatory neurons:

```{r, fig.width=7, fig.height=5, fig.fullwidth=TRUE}

FeaturePlot(seur_comb, features = c("FEZF2",
                                    "RORB", 
                                    "THEMIS"))

```

Astrocytes
```{r, fig.width=7, fig.height=15, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("GJA1",
                                    "AQP4", 
                                    "GLUL", 
                                    "SOX9", 
                                    "NDRG2", 
                                    "GFAP", 
                                    "ALDH1A1", 
                                    "ALDH1L1", 
                                    "VIM", 
                                    "APOE", 
                                    "FGFR3"),
            ncol = 2)


```

Add Astrocytes
```{r, fig.width=7, fig.height=35, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c( "HPSE2", 
                                     "TENM2",  
                                     "ZNF98",
                                     "CCDC85A", 
                                     "SLC38A1",
                                     "GALNT15",
                                     "BNC2",
                                     "CFAP299",
                                     "SPAG17",
                                     "DTHD1",
                                     "ADGB",
                                     "UAP1",
                                     "SPOCD1",
                                     "CLCF1",
                                     "CDH19",
                                     "SORCS1",
                                     "SAMHD1",
                                     "NRGN",
                                     "CAMK2A",
                                     "THY1",
                                     "ENC1",
                                     "SYT1",
                                     "CALM3",
                                     "ST18",
                                     "CTNNA3",
                                     "APLNR"))
            
         

```


Astrocyte markers as described in Zeisel et al. 2018 in the mouse 
for telencephalon and non-telencephalon astrocytes


```{r, fig.width=7, fig.height=13, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c( "AGT", 
                                     "MFGE8",  
                                     "SLC6A11",
                                     "SLC6A9", 
                                     "GDF10",
                                     "ISLR",
                                     "GFAP",
                                     "AQP4"))
            
         

```

Endothelial cells and pericytes
```{r, fig.width=7, fig.height=10, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c( "CLDN5",   
                                     "ICAM2",
                                     "PDGFRB", 
                                     "NOTCH3", 
                                     "VWF",
                                     "FLT1",
                                     "MECOM"), 
            ncol = 2)
            
         

```






Stromal

```{r, fig.width=7, fig.height=8, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c( "LAMB1" , 
                                     "HSPG2", 
                                     "COL4A1", 
                                     "FN1", 
                                     "LAMA2"),
            ncol = 2)




```

Microglia and macrophages
```{r, fig.width=7, fig.height=13, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c( "CD74", 
                                     "SPI1", 
                                     "MRC1", 
                                     "TMEM119", 
                                     "CX3CR1", 
                                     "AIF1",
                                     "P2RY12",
                                     "C1QC",
                                     "C1QA"),
            ncol = 2)




```
Add microglia & macrophages

```{r, fig.width=7, fig.height=13, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("IPCEF1", 
                                    "NAV3", 
                                    "NAMPT", 
                                    "FOSL2",
                                    "IPCEF1", 
                                    "NAV3", 
                                    "IL1RAPL1", 
                                    "FRMD5",
                                    "MAGI2"),
            ncol = 2)




```



OPCs
```{r, fig.width=7, fig.height=13, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("PDGFRA", 
                                    "CSPG4", 
                                    "GPR17", 
                                    "PTPRZ1",
                                    "OLIG1", 
                                    "OLIG2", 
                                    "PCDH15", 
                                    "PTGDS",
                                    "BCAN"),
            ncol = 2)




```

Add OPC:


```{r, fig.width=7, fig.height=5, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("CABLES1", 
                                    "GFRA1", 
                                    "LINC01965"), 
            ncol = 2)


```
Add COP:


```{r, fig.width=7, fig.height=18, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("ETV1", 
                                    "CHST9", 
                                    "MYT1", 
                                    "TENM2", 
                                    "CAMK2A", 
                                    "KCNQ5", 
                                    "SEMA5B",
                                    "SYT1",
                                    "GPR17",
                                    "BMPER",
                                    "EPHB1",
                                    "ARHGAP24",
                                    "DOCK8",
                                    "APBB1IP"), 
            ncol = 2)


```



Oligodendrocytes
```{r, fig.width=7, fig.height=10, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("PLP1", 
                                    "CNP", 
                                    "MAG", 
                                    "MOG", 
                                    "MOBP", 
                                    "MBP", 
                                    "SOX10" ), 
            ncol = 2)


```


Additional oligodendrocytes
```{r, fig.width=7, fig.height=18, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("SNAP25", 
                                    "CNTN1", 
                                    "FRY", 
                                    "PLXDC2", 
                                    "DYSF", 
                                    "PTPRM", 
                                    "FOS",
                                    "VIM",
                                    "JUNB",
                                    "AFF3",
                                    "KANK4",
                                    "FSTL5",
                                    "SGCZ",
                                    "MDGA2"), 
            ncol = 2)


```



```{r, fig.width=12, fig.height=30, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("SNAP25", 
                                    "STMN2", 
                                    "RBFOX3",
                                    "GABRB2", 
                                    "GAD1", 
                                    "GAD2", 
                                    "SLC32A1", 
                                    "SATB2", 
                                    "SLC12A6", 
                                    "SLC17A7", 
                                    "GJA1",
                                    "AQP4", 
                                    "GLUL", 
                                    "SOX9", 
                                    "NDRG2", 
                                    "GFAP", 
                                    "ALDH1A1", 
                                    "ALDH1L1", 
                                    "VIM",  
                                    "CLDN5", 
                                    "VTN",  
                                    "ICAM2",
                                    "PDGFRB", 
                                    "NOTCH3", 
                                    "CD74", 
                                    "SPI1", 
                                    "MRC1", 
                                    "TMEM119", 
                                    "CX3CR1", 
                                    "AIF1", 
                                    "PDGFRA", 
                                    "CSPG4", 
                                    "GPR17", 
                                    "PTPRZ1",
                                    "OLIG1", 
                                    "OLIG2", 
                                    "PCDH15", 
                                    "PTGDS", 
                                    "PLP1", 
                                    "CNP", 
                                    "MAG", 
                                    "MOG", 
                                    "MOBP",
                                    "MBP", 
                                    "SOX10"), 
            ncol = 4)


```

```{r, fig.width=12, fig.height=12, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("SNAP25", 
                                    "GABRB2",
                                    "GJA1",
                                    "AQP4", 
                                    "CLDN5",
                                    "ICAM2",
                                    "PDGFRB", 
                                    "NOTCH3", 
                                    "CD74", 
                                    "AIF1", 
                                    "PDGFRA", 
                                    "CSPG4", 
                                    "GPR17", 
                                    "PTPRZ1", 
                                    "PLP1", 
                                    "CNP"), 
            ncol = 4)


```


```{r, fig.width=8, fig.height=6, fig.fullwidth=TRUE}

DimPlot(seur_comb, label = T, cols = mycoloursP)

```




Look for markers previously found interesting in our analyses:

```{r, fig.width=12, fig.height=12, fig.fullwidth=TRUE}
FeaturePlot(seur_comb, features = c("RBFOX1",
                                    "SPARC",
                                    "OPALIN",
                                    "PAX3",
                                    "NELL1",
                                    "KLK6"))


```