---
title: "Nadine nad_ol"
author: "Luise A. Seeker"
date: "26/02/2021"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 4
    theme: united
---

# Introduction

The purpose of this script is to find variable genes, cluster the oligodendroglia
dataset at different resolutions, find some rough markers for cluster
stability/ purity that may help to decide which clusters to use, and find 
marker genes for those clusters. 

It is a long and repetitive script that could be written more elegantly, but
this work is very explorative, manual and depending on assessment of visual 
output. Making it prettier now that I know the resolutiions I would like to test
tempts me, but does not bring me further in the analysis. So therefore I decided
to keep the script for now as it is. 




# Load libraries

```{r, echo = F}
# general
library(here)
library(tibble)
library(dplyr)


# plotting
library(ggplot2)
library(ggdendro)
library(ggsci)
library(gridExtra)
library(viridis) 
library(scales)
library(pheatmap)
library(dendextend)

# stats
library(NMF)
library(limma)
library(StatMeasures)

# other
library(zoo)
library(philentropy)
library(bluster)

# sc Analysis
library(scran)
library(scater)
library(SingleCellExperiment)
library(EnhancedVolcano)
library(Seurat)
library(clustree)

#For monocle pseudotime:
library(monocle3)
library(SeuratWrappers)





```


# Pick colour pallets

```{r, echo = F}
mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)

mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)

show_col(mycoloursP, labels =F)
```


# Read in dataset

```{r, echo = F}

nad_ol <- readRDS("/Users/lseeker/Documents/Work/HumanCellAtlas/srt_oligos_Nadine/srt_oligos_and_opcs_LS.RDS")

```



# Non-linear dimensional reduction
```{r}


FeaturePlot(nad_ol, features = "PAX3")
FeaturePlot(nad_ol, features = "PLP1")
FeaturePlot(nad_ol, features = "MAG")
FeaturePlot(nad_ol, features = "MOG")
FeaturePlot(nad_ol, features = "MBP")
FeaturePlot(nad_ol, features = "OPALIN")
FeaturePlot(nad_ol, features = "SPARC")
FeaturePlot(nad_ol, features = "SPARCL1")
FeaturePlot(nad_ol, features = "PCDH15")
FeaturePlot(nad_ol, features = "PDGFRA")
FeaturePlot(nad_ol, features = "nCount_RNA")
FeaturePlot(nad_ol, features = "nFeature_RNA")

```





# Test for batch effect etc

```{r, fig.width = 10, fig.height = 8}

DimPlot(nad_ol, 
        reduction = "umap", 
        cols= mycoloursP[6:40], 
        group.by = "Tissue",
        split.by = "caseNO", 
        pt.size = 0.8, 
        ncol = 5)
```

```{r, fig.width = 10, fig.height = 8}

DimPlot(nad_ol, 
        reduction = "umap", 
        cols= mycoloursP[6:40], 
        split.by = "uniq_id", 
        group.by = "Tissue", 
        pt.size = 0.8, 
        ncol = 5)
```


# Test different clustering resolutions
## 0.5

```{r}

DimPlot(nad_ol, reduction = "umap", 
        cols= mycoloursP, 
        group.by = "RNA_snn_res.0.5",
        label = TRUE)

```

# Cerebellum separately
```{r}

Idents(nad_ol) <- "Tissue"
cb_srt <- subset(nad_ol, idents = "CB")
cb_srt <- FindVariableFeatures(cb_srt, 
                               selection.method = "vst", 
                               nfeatures = 2000)
all_genes_cb_srt <- rownames(cb_srt)
cb_srt <- ScaleData(cb_srt, features= all_genes_cb_srt)
cb_srt <- RunPCA(cb_srt, features = VariableFeatures(object = cb_srt))
cb_srt <- FindNeighbors(cb_srt, dims = 1:10)
cb_srt <- FindClusters(cb_srt, resolution = 0.5)
cb_srt <- RunUMAP(cb_srt, dims = 1:10)
DimPlot(cb_srt, reduction = "umap", cols= mycoloursP[4:40])

```


```{r}
FeaturePlot(cb_srt, features = "OPALIN")

```

```{r}
FeaturePlot(cb_srt, features = "SPARC")

```

# Cervical spinal cord separately
```{r}

Idents(nad_ol) <- "Tissue"
csc_srt <- subset(nad_ol, idents = "CSC")
csc_srt <- FindVariableFeatures(csc_srt, 
                               selection.method = "vst", 
                               nfeatures = 2000)
all_genes_csc_srt <- rownames(csc_srt)
csc_srt <- ScaleData(csc_srt, features= all_genes_csc_srt)
csc_srt <- RunPCA(csc_srt, features = VariableFeatures(object = csc_srt))
csc_srt <- FindNeighbors(csc_srt, dims = 1:7)
csc_srt <- FindClusters(csc_srt, resolution = 0.7)
csc_srt <- RunUMAP(csc_srt, dims = 1:7)
DimPlot(csc_srt, reduction = "umap", cols= mycoloursP[4:40], label = TRUE)

```





```{r}
FeaturePlot(csc_srt, features = "RBFOX1")

```


```{r}
FeaturePlot(csc_srt, features = "SPARC")

```


```{r}
FeaturePlot(csc_srt, features = "OPALIN")

```


```{r}
FeaturePlot(csc_srt, features = "PLP1")

```


```{r}
FeaturePlot(csc_srt, features = "PDGFRA")

```


```{r}
FeaturePlot(csc_srt, features = "SPARCL1")

```
```{r}
FeaturePlot(csc_srt, features = "KLK6")

```


# BA4 spinal cord separately
```{r}

Idents(nad_ol) <- "Tissue"
ba4_srt <- subset(nad_ol, idents = "BA4")
ba4_srt <- FindVariableFeatures(ba4_srt, 
                               selection.method = "vst", 
                               nfeatures = 2000)
all_genes_ba4_srt <- rownames(ba4_srt)
ba4_srt <- ScaleData(ba4_srt, features= all_genes_ba4_srt)
ba4_srt <- RunPCA(ba4_srt, features = VariableFeatures(object = ba4_srt))
ba4_srt <- FindNeighbors(ba4_srt, dims = 1:10)
ba4_srt <- FindClusters(ba4_srt, resolution = 0.5)
ba4_srt <- RunUMAP(ba4_srt, dims = 1:10)
DimPlot(ba4_srt, reduction = "umap", cols= mycoloursP[4:40])


```

```{r}

DimPlot(ba4_srt, reduction = "umap", cols= mycoloursP[4:40],
        group.by = "caseNO")


```

```{r}

DimPlot(ba4_srt, reduction = "umap", cols= mycoloursP[4:40],
        group.by = "caseNO", split.by = "caseNO", ncol = 4)

```

```{r}
clu_mark <- FindMarkers(ba4_srt, ident.1 = 6, ident.2 = c(1,2,3,4,5,7), only.pos = TRUE)
clu_mark
```


```{r}
FeaturePlot(ba4_srt, features = "AL139260.1")

```

```{r}

FeaturePlot(ba4_srt, features = "ABHD3")

```

```{r}

FeaturePlot(ba4_srt, features = "SNAP25")

```

```{r}

FeaturePlot(ba4_srt, features = "RBFOX3")

```

```{r}

FeaturePlot(ba4_srt, features = "DNAJB1")
FeaturePlot(ba4_srt, features = "BACH2")
```


```{r}
FeaturePlot(ba4_srt, features = "FOS")

```

```{r}
FeaturePlot(ba4_srt, features = "FOS")

```

```{r}
FeaturePlot(ba4_srt, features = "OPALIN")

```

```{r}
FeaturePlot(ba4_srt, features = "SPARC")

```

```{r}
FeaturePlot(ba4_srt, features = "SPARCL1")

```
```{r}
FeaturePlot(ba4_srt, features = "RBFOX1")

```

```{r}
FeaturePlot(ba4_srt, features = "GFAP")

```


```{r}
FeaturePlot(ba4_srt, features = "PLP1")

```


```{r}
FeaturePlot(ba4_srt, features = "PLP1")

```


```{r}
FeaturePlot(ba4_srt, features = "PDGFRA")

```


```{r}
FeaturePlot(ba4_srt, features = "MAG")

```
```{r}
FeaturePlot(ba4_srt, features = "KLK6")

```

```{r}
merge_ol <- merge(csc_srt, y = c(cb_srt, ba4_srt), add.cell.ids = c("csc", "cb", "ba4"))

merge_ol <- FindVariableFeatures(merge_ol, 
                               selection.method = "vst", 
                               nfeatures = 2000)
all_genes_merge_ol <- rownames(merge_ol)
merge_ol <- ScaleData(merge_ol, features= all_genes_merge_ol)
merge_ol <- RunPCA(merge_ol, features = VariableFeatures(object = merge_ol))
merge_ol <- FindNeighbors(merge_ol, dims = 1:10)
merge_ol <- FindClusters(merge_ol, resolution = 0.5)
merge_ol <- RunUMAP(merge_ol, dims = 1:10)
DimPlot(merge_ol, reduction = "umap", cols= mycoloursP[4:40])

FeaturePlot(merge_ol, features = "OPALIN")
FeaturePlot(merge_ol, features = "RBFOX1")
FeaturePlot(merge_ol, features = "SPARC")
FeaturePlot(merge_ol, features = "SPARCL1")
FeaturePlot(merge_ol, features = "NELL1")
FeaturePlot(merge_ol, features = "PAX3")

```







`