---
title: "Cholesterol_biosythesis_genes"
author: "Luise A. Seeker"
date: "01/07/2021"
output: html_document
---
# Introduction

I watched Professor Terry Wood from Rutgers University presenting a comparison 
of murine single cell RNAseq data from  brain and spinal cord at Euroglia 2021. 
She showed that in mice the most obvious difference between those tissues was
that spinal cord oligodendroglia expressed more cholesterol biosynthesis genes. 
This makes sense considering that they produce longer myelin sheaths and thus
more myelin per cell and myelin production requires cholesterol. 

Cholesterol biosynthesis genes have never particularly shown up in my 
differential gene expression analyses, and still, I think it's a good idea to at
least quickly check. 

```{r}
library(Seurat)
library(ggplot2)
library(here)
```

```{r}

nad_ol <- nad_ol <- readRDS(here("data", 
                                 "single_nuc_data", 
                                 "oligodendroglia",
                                 "srt_oligos_and_opcs_LS.RDS"))
```



```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "ACAT2", split.by = "Tissue")
```


```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "HMGCS1", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "HMGCR", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "PMVK", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "MVD", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "IDI1", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "FDPS", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "FDFT1", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "SQLE", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "DHCR24", split.by = "Tissue")
```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "MSMO1", split.by = "Tissue")

```

```{r, fig.width = 9, fig.height = 2.5}
FeaturePlot(nad_ol, features = "SC5D", split.by = "Tissue")

```

# Conclusion

It does not look like cholesterol genes are particularly enriched in the spinal 
cord oligodendroglia. The observed difference to the mouse data may be due to 
species differences or due to age differences. Most myelination happens soon post
partum, but myelination at adult ages, particularly in the absence of injury/
disease is less pronounced. There may be a tissue difference at earlier ages
that is not possible to observe anymore at adult ages. 

# SessionInfo


```{r}
sessionInfo()
```