---
title: "Milo"
author: "Luise A. Seeker"
date: "22/10/2021"
output: html_document
---

```{r}
library(Seurat)
library(here)
library(miloR)
library(dplyr)
library(statmod)
library(SingleCellExperiment)
library(scater)
library(patchwork)
library(scran)
```

```{r}
seur <- readRDS(here("data", 
                       "single_nuc_data", 
                       "oligodendroglia",
                       "srt_oligos_and_opcs_LS.RDS"))

```

```{r}

plot_levels <- c("Oligo_F",
                                                          "Oligo_E",
                                                          "Oligo_D",
                                                          "Oligo_C",
                                                          "Oligo_B",
                                                          "Oligo_A",
                                                          "COP_C",
                                                          "COP_B",
                                                          "COP_A",
                                                          "OPC_B",
                                                          "OPC_A")
```


```{r}
milo_meta <- seur@meta.data
milo_obj <- Milo(as.SingleCellExperiment(seur))
milo_obj

```


Build a graph and neighbourhoods.
```{r}
milo_obj <- buildGraph(milo_obj, k=30, d=30, reduced.dim = "PCA" )
milo_obj <- makeNhoods(milo_obj, k=30, d=30, refined=TRUE, prop=0.2, reduced_dims = "PCA" )

plotNhoodSizeHist(milo_obj)
```
Ideally, histogram should be peaking at 50 -100. Or the average neighbourhood size
should be  over 5 x N_samples.Increasing k and/or
prop during makeNhoods will shift distribution to the right. 


# Calculate distances, count cells according to an experimental design and perform DA testing.

```{r}
#milo_obj <- calcNhoodDistance(milo_obj, d=30)
milo_obj <- countCells(milo_obj, samples="uniq_id", meta.data=milo_meta)
head(nhoodCounts(milo_obj))
```

```{r}
milo_design <- data.frame(colData(milo_obj))[,c("uniq_id", "Tissue", "gender",
                                                "AgeGroup", "caseNO")]

## Convert batch info from integer to factor
milo_design <- distinct(milo_design)
rownames(milo_design) <- milo_design$uniq_id

milo_design

```



```{r}
milo_obj <- calcNhoodDistance(milo_obj, d=30, reduced.dim = "PCA")

```


Testing


```{r}

da_results <- testNhoods(milo_obj, design = ~ Tissue, design.df = milo_design)
head(da_results)
```


```{r}
da_results %>%
  arrange(SpatialFDR) %>%
  head() 
```
Inspecting DA results

```{r}
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)

```

```{r}
ggplot(da_results, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```

```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Plot single-cell UMAP
umap_pl <- plotReducedDim(milo_obj, dimred = "UMAP", colour_by="Tissue", 
                          text_by = "ol_clusters_named", 
                          text_size = 3, point_size=0.5) +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results, layout="UMAP",alpha=0.1) 
  
umap_pl + nh_graph_pl +
  plot_layout(guides="collect")

```


```{r}
da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "ol_clusters_named")
head(da_results)

```

```{r}

ggplot(da_results, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```


```{r}
da_results$clust <- ifelse(da_results$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results$ol_clusters_named_fraction)

plotDAbeeswarm(da_results, group.by = "ol_clusters_named")
```

# Pairwise tissue
CB vs BA4 
```{r}
Idents(seur) <- "Tissue"
cb_ba4_srt <- subset(seur, idents = c("CB", "BA4"))
milo_meta_cb_ba4 <- cb_ba4_srt@meta.data
milo_obj <- Milo(as.SingleCellExperiment(cb_ba4_srt))
milo_obj

```


Build a graph and neighbourhoods.
```{r}
milo_obj <- buildGraph(milo_obj, k=30, d=30, reduced.dim = "PCA" )
milo_obj <- makeNhoods(milo_obj, k=30, d=30, refined=TRUE, prop=0.2, reduced_dims = "PCA" )

plotNhoodSizeHist(milo_obj)
```
Ideally, histogram should be peaking at 50 -100. Or the average neighbourhood size
should be  over 5 x N_samples.Increasing k and/or
prop during makeNhoods will shift distribution to the right. 


# Calculate distances, count cells according to an experimental design and perform DA testing.

```{r}
#milo_obj <- calcNhoodDistance(milo_obj, d=30)
milo_obj <- countCells(milo_obj, samples="uniq_id", meta.data=milo_meta_cb_ba4)
head(nhoodCounts(milo_obj))
```

```{r}
milo_design <- data.frame(colData(milo_obj))[,c("uniq_id", "Tissue", "gender",
                                                "AgeGroup", "caseNO")]

## Convert batch info from integer to factor
milo_design <- distinct(milo_design)
rownames(milo_design) <- milo_design$uniq_id

milo_design

```



```{r}
set.seed(1001)
milo_obj <- calcNhoodDistance(milo_obj, d=30, reduced.dim = "PCA")

```


```{r}

set.seed(1001)


da_results_cb_vs_ba4 <- testNhoods(milo_obj, design = ~ 0 + Tissue, design.df = milo_design,
                         model.contrasts = c("TissueCB - TissueBA4"))


head(da_results_cb_vs_ba4)
```


```{r}
da_results_cb_vs_ba4 %>%
  arrange(SpatialFDR) %>%
  head() 
```
Inspecting DA results

```{r}
ggplot(da_results_cb_vs_ba4, aes(PValue)) + geom_histogram(bins=50)

```

```{r}
ggplot(da_results_cb_vs_ba4, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```

```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Plot single-cell UMAP
umap_pl <- plotReducedDim(milo_obj, dimred = "UMAP", colour_by="Tissue", 
                          text_by = "ol_clusters_named", 
                          text_size = 3, point_size=0.5) +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results_cb_vs_ba4, layout="UMAP",alpha=0.1) 
  
umap_cb_ba4 <- umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
umap_cb_ba4
```

```{r}
milo_save_dir <- here("outs",
                 "Milo_compositional_analysis",
                 "oligos_only",
                 "umap_plots")

dir.create(milo_save_dir, recursive = T)

                       
pdf(file = here(milo_save_dir, "Tissue_CB_vs_BA4_noCSC"),   
              width = 13, 
              height = 7) 
print(umap_cb_ba4)
dev.off()

```


```{r}
da_results_cb_vs_ba4 <- annotateNhoods(milo_obj, 
                                       da_results_cb_vs_ba4, 
                                       coldata_col = "ol_clusters_named")
head(da_results_cb_vs_ba4)

```

```{r}

ggplot(da_results_cb_vs_ba4, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```


```{r}
da_results_cb_vs_ba4$clust <- ifelse(da_results_cb_vs_ba4$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results_cb_vs_ba4$ol_clusters_named_fraction)

da_results_cb_vs_ba4$ol_clusters_named <- factor(da_results_cb_vs_ba4$ol_clusters_named,
                                            levels = plot_levels)

cb_ba4 <- plotDAbeeswarm(da_results_cb_vs_ba4, group.by = "ol_clusters_named")

cb_ba4
```

```{r}
milo_dir_bees <- here("outs",
                 "Milo_compositional_analysis",
                 "oligos_only",
                 "beeswarm_plots")

dir.create(milo_dir_bees, recursive = T)

                       
pdf(file = here(milo_dir_bees, "Tissue_CB_vs_BA4_noCSC"),   
              width = 7, 
              height = 13) 
print(cb_ba4)
dev.off()

```



# Pairwise tissue

CB vs CSC
```{r}
Idents(nad_ol) <- "Tissue"
cb_csc_srt <- subset(nad_ol, idents = c("CSC", "CB"))

milo_meta_cb_csc <- cb_csc_srt@meta.data
milo_obj <- Milo(as.SingleCellExperiment(cb_csc_srt))
milo_obj
```

Build a graph and neighbourhoods.
```{r}
milo_obj <- buildGraph(milo_obj, k=30, d=30, reduced.dim = "PCA" )
milo_obj <- makeNhoods(milo_obj, k=30, d=30, refined=TRUE, prop=0.2, reduced_dims = "PCA" )

plotNhoodSizeHist(milo_obj)
```


# Calculate distances, count cells according to an experimental design and perform DA testing.

```{r}
#milo_obj <- calcNhoodDistance(milo_obj, d=30)
milo_obj <- countCells(milo_obj, samples="uniq_id", meta.data=milo_meta_cb_csc)
head(nhoodCounts(milo_obj))
```

```{r}
milo_design <- data.frame(colData(milo_obj))[,c("uniq_id", "Tissue", "gender",
                                                "AgeGroup", "caseNO")]

## Convert batch info from integer to factor
milo_design <- distinct(milo_design)
rownames(milo_design) <- milo_design$uniq_id

milo_design

```



```{r}
set.seed(1001)
milo_obj <- calcNhoodDistance(milo_obj, d=30, reduced.dim = "PCA")

```


```{r}

set.seed(1001)


da_results_cb_vs_csc <- testNhoods(milo_obj, design = ~ 0 + Tissue, 
                                   design.df = milo_design,
                         model.contrasts = c("TissueCB - TissueCSC"))
head(da_results_cb_vs_csc)
```


```{r}
da_results_cb_vs_csc %>%
  arrange(SpatialFDR) %>%
  head() 
```
Inspecting DA results

```{r}
ggplot(da_results_cb_vs_csc, aes(PValue)) + geom_histogram(bins=50)

```

```{r}
ggplot(da_results_cb_vs_csc, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```

```{r}
milo_obj <- buildNhoodGraph(milo_obj) 
## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results_cb_vs_csc, layout="UMAP",alpha=0.1) 
  
umap_cb_csc <- umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
umap_cb_csc
```


```{r}
                       
pdf(file = here(milo_save_dir, "Tissue_CB_vs_CSC_noBA4"),   
              width = 13, 
              height = 7) 
print(umap_cb_csc)
dev.off()

```


```{r}
da_results_cb_vs_csc <- annotateNhoods(milo_obj, 
                                       da_results_cb_vs_csc, 
                                       coldata_col = "ol_clusters_named")
head(da_results_cb_vs_csc)

```

```{r}

ggplot(da_results_cb_vs_csc, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```


```{r}
da_results_cb_vs_csc$clust <- ifelse(da_results_cb_vs_csc$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results_cb_vs_csc$ol_clusters_named_fraction)

da_results_cb_vs_csc$ol_clusters_named <- factor(da_results_cb_vs_csc$ol_clusters_named,
                                            levels = plot_levels)

cb_csc <- plotDAbeeswarm(da_results_cb_vs_csc, group.by = "ol_clusters_named")

cb_csc
```

```{r}

                       
pdf(file = here(milo_dir_bees, "Tissue_CB_vs_CSC"),   
              width = 7, 
              height = 13) 
print(cb_csc)
dev.off()

```


# Pairwise tissue

CSC vs BA4

```{r}
Idents(seur) <- "Tissue"
csc_ba4_srt <- subset(seur, idents = c("CSC", "BA4"))
milo_meta_csc_ba4 <- csc_ba4_srt@meta.data
milo_obj <- Milo(as.SingleCellExperiment(csc_ba4_srt))
milo_obj

```


Build a graph and neighbourhoods.
```{r}
milo_obj <- buildGraph(milo_obj, k=30, d=30, reduced.dim = "PCA" )
milo_obj <- makeNhoods(milo_obj, k=30, d=30, refined=TRUE, prop=0.2, reduced_dims = "PCA" )

plotNhoodSizeHist(milo_obj)
```
Ideally, histogram should be peaking at 50 -100. Or the average neighbourhood size
should be  over 5 x N_samples.Increasing k and/or
prop during makeNhoods will shift distribution to the right. 


# Calculate distances, count cells according to an experimental design and perform DA testing.

```{r}
#milo_obj <- calcNhoodDistance(milo_obj, d=30)
milo_obj <- countCells(milo_obj, samples="uniq_id", meta.data=milo_meta_csc_ba4)
head(nhoodCounts(milo_obj))
```

```{r}
milo_design <- data.frame(colData(milo_obj))[,c("uniq_id", "Tissue", "gender",
                                                "AgeGroup", "caseNO")]

## Convert batch info from integer to factor
milo_design <- distinct(milo_design)
rownames(milo_design) <- milo_design$uniq_id

milo_design

```



```{r}
set.seed(1001)
milo_obj <- calcNhoodDistance(milo_obj, d=30, reduced.dim = "PCA")

```

```{r}


da_results_csc_vs_ba4 <- testNhoods(milo_obj, design = ~ 0 + Tissue, design.df = milo_design,
                         model.contrasts = c("TissueCSC - TissueBA4"))

head(da_results_csc_vs_ba4)
```


```{r}
da_results_csc_vs_ba4 %>%
  arrange(SpatialFDR) %>%
  head() 
```
Inspecting DA results

```{r}
ggplot(da_results_csc_vs_ba4, aes(PValue)) + geom_histogram(bins=50)

```

```{r}
ggplot(da_results_csc_vs_ba4, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```

```{r}
milo_obj <- buildNhoodGraph(milo_obj)
## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results_csc_vs_ba4, layout="UMAP",alpha=0.1) 
  
umap_csc_ba4 <- umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
umap_csc_ba4
```


```{r}
                       
pdf(file = here(milo_save_dir, "Tissue_CSC_vs_BA4_noCB"),   
              width = 13, 
              height = 7) 
print(umap_csc_ba4)
dev.off()

```


```{r}
da_results_csc_vs_ba4 <- annotateNhoods(milo_obj, 
                                       da_results_csc_vs_ba4, 
                                       coldata_col = "ol_clusters_named")
head(da_results_csc_vs_ba4)

```

```{r}

ggplot(da_results_csc_vs_ba4, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```


```{r}
da_results_csc_vs_ba4$clust <- ifelse(da_results_csc_vs_ba4$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results_csc_vs_ba4$ol_clusters_named_fraction)

da_results_csc_vs_ba4$ol_clusters_named <- factor(da_results_csc_vs_ba4$ol_clusters_named,
                                            levels = plot_levels)

csc_ba4 <- plotDAbeeswarm(da_results_csc_vs_ba4, group.by = "ol_clusters_named")

csc_ba4
```

```{r}

grid.arrange(csc_ba4, cb_ba4, cb_csc, ncol = 3)

```


```{r}

                       
pdf(file = here(milo_dir_bees, "Tissue_CSC_vs_BA4_noCB"),   
              width = 7, 
              height = 13) 
print(csc_ba4)
dev.off()

```


###




# Finding markers of DA populations

## Automatic grouping of neighborhoods

```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Find groups
da_results <- groupNhoods(milo_obj, da_results, max.lfc.delta = 2)
head(da_results)

```

```{r}

plotNhoodGroups(milo_obj, da_results, layout="UMAP") 
```


```{r}

plotDAbeeswarm(da_results, "NhoodGroup")
```


```{r, fig.height = 5, fig.width = 7}
plotDAbeeswarm(groupNhoods(milo_obj, da_results, max.lfc.delta = 0.5), 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=0.5")

```

```{r, fig.height = 5, fig.width = 7}
plotDAbeeswarm(groupNhoods(milo_obj, da_results, max.lfc.delta = 0.2), 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=0.2")

```

```{r}
plotDAbeeswarm(groupNhoods(milo_obj, da_results, max.lfc.delta = 1) , 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=1")


```

```{r}
plotDAbeeswarm(groupNhoods(milo_obj, da_results, max.lfc.delta = 1, 
                           overlap=3), group.by = "NhoodGroup") + 
  ggtitle("overlap=3")

```


```{r}
set.seed(42)
da_results <- groupNhoods(milo_obj, da_results, max.lfc.delta = 1, overlap=3)
plotNhoodGroups(milo_obj, da_results, layout="UMAP")

```



```{r}
plotDAbeeswarm(da_results, group.by = "NhoodGroup")

```
# Finding gene signatures for neighbourhoods

```{r, eval = FALSE}
keep_rows <- rowSums(logcounts(milo_obj)) != 0
milo_obj <- milo_obj[keep_rows, ]

## Find HVGs
dec <- modelGeneVar(milo_obj)
hvgs <- getTopHVGs(dec, n=2000)
head(hvgs)
```



```{r, eval= F}
nhood_markers <- findNhoodGroupMarkers(milo_obj, 
                                       da_results, 
                                       subset.row = hvgs, 
                                       aggregate.samples = TRUE, 
                                       sample_col = "uniq_id")

head(nhood_markers)

save_dir <- here("outs",
                 "Milo_compositional_analysis",
                 "Milo_neighbourhood_mark")
dir.create(save_dir)
write.csv(nhood_markers, here(save_dir,
                              "Milo_neighbour_markers.csv"))
```


# Age 

Testing


```{r}
da_results_age <- testNhoods(milo_obj, design = ~ Tissue + gender + AgeGroup, 
                             design.df = milo_design)
head(da_results_age)
```


```{r}
da_results_age %>%
  arrange(SpatialFDR) %>%
  head() 
```
Inspecting DA results

```{r}
ggplot(da_results_age, aes(PValue)) + geom_histogram(bins=50)

```

```{r}
ggplot(da_results_age, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```

```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Plot single-cell UMAP
umap_pl <- plotReducedDim(milo_obj, dimred = "UMAP", colour_by="AgeGroup", 
                          text_by = "ol_clusters_named", 
                          text_size = 3, point_size=0.5) +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results_age, layout="UMAP",alpha=0.1) 
  
age_umap <-umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
age_umap
```

```{r}
da_results_age <- annotateNhoods(milo_obj, 
                                       da_results_age, 
                                       coldata_col = "ol_clusters_named")
head(da_results_age)

```

```{r}

ggplot(da_results_age, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```

```{r}
da_results_age$clust <- ifelse(da_results_age$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results_age$ol_clusters_named_fraction)

da_results_age$ol_clusters_named <- factor(da_results_age$ol_clusters_named,
                                            levels = plot_levels)


plotDAbeeswarm(da_results_age, group.by = "ol_clusters_named")

```

```{r}
                       
pdf(file = here(milo_save_dir, "Age_corr_for_tissue_sex.pdf"),   
              width = 13, 
              height = 7) 
print(age_umap)
dev.off()

```


```{r}
da_results_age <- annotateNhoods(milo_obj, da_results_age, coldata_col = "ol_clusters_named")
head(da_results_age)

```

```{r}

da_results_age$ol_clusters_named <- factor(da_results_age$ol_clusters_named, 
                                      levels = plot_levels)

ggplot(da_results_age, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```




```{r}
da_results_age$clust <- ifelse(da_results_age$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results_age$ol_clusters_named_fraction)

age_bees <- plotDAbeeswarm(da_results_age, group.by = "ol_clusters_named")
```


```{r}
milo_dir_bees <- here("outs",
                 "Milo_compositional_analysis",
                 "beeswarm_plots")

                       
pdf(file = here(milo_dir_bees, "Age_correct_for_tissue_sex.pdf"),   
              width = 7, 
              height = 13) 
print(age_bees)
dev.off()

```



# Finding markers of DA populations

## Automatic grouping of neighborhoods

```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Find groups
da_results_age <- groupNhoods(milo_obj, da_results_age, max.lfc.delta = 2)
head(da_results_age)

```

```{r}

plotNhoodGroups(milo_obj, da_results_age, layout="UMAP") 
```


```{r}

plotDAbeeswarm(da_results_age, "NhoodGroup")
```


```{r, fig.height = 5, fig.width = 7}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_age, max.lfc.delta = 0.5), 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=0.5")

```

```{r, fig.height = 5, fig.width = 7}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_age, max.lfc.delta = 0.2), 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=0.2")

```

```{r}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_age, max.lfc.delta = 1) , 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=1")


```

```{r}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_age, max.lfc.delta = 1, 
                           overlap=3), group.by = "NhoodGroup") + 
  ggtitle("overlap=3")

```


```{r}
set.seed(42)
da_results_age <- groupNhoods(milo_obj, da_results_age, max.lfc.delta = 1, overlap=3)
plotNhoodGroups(milo_obj, da_results_age, layout="UMAP")

```



```{r}
plotDAbeeswarm(da_results_age, group.by = "NhoodGroup")

```
# Finding gene signatures for neighbourhoods

```{r, eval = FALSE}
keep_rows <- rowSums(logcounts(milo_obj)) != 0
milo_obj <- milo_obj[keep_rows, ]

## Find HVGs
dec <- modelGeneVar(milo_obj)
hvgs <- getTopHVGs(dec, n=2000)
head(hvgs)
```



```{r, eval= F}
nhood_markers <- findNhoodGroupMarkers(milo_obj, 
                                       da_results_age, 
                                       subset.row = hvgs, 
                                       aggregate.samples = TRUE, 
                                       sample_col = "uniq_id")

head(nhood_markers)

save_dir <- here("outs",
                 "Milo_compositional_analysis",
                 "Milo_neighbourhood_mark")
dir.create(save_dir)
write.csv(nhood_markers, here(save_dir,
                              "Milo_neighbour_markers.csv"))
```



# Sex

Testing


```{r}

da_results_sex <- testNhoods(milo_obj, design = ~ Tissue+ AgeGroup+ gender, 
                             design.df = milo_design)
head(da_results_sex)
```


```{r}
da_results_sex %>%
  arrange(SpatialFDR) %>%
  head() 
```
Inspecting DA results

```{r}
ggplot(da_results_sex, aes(PValue)) + geom_histogram(bins=50)

```

```{r}
ggplot(da_results_sex, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```

```{r}

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results_sex, layout="UMAP",alpha=0.1) 
  
sex_umap <- umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
sex_umap
```




```{r}
                       
pdf(file = here(milo_save_dir, "Sex_corr_for_tissue_age.pdf"),   
              width = 13, 
              height = 7) 
print(sex_umap)
dev.off()

```


```{r}
da_results_sex <- annotateNhoods(milo_obj, da_results_sex, coldata_col = "ol_clusters_named")
head(da_results_sex)

```

```{r}
da_results_sex$ol_clusters_named <- factor(da_results_sex$ol_clusters_named, 
                                      levels = plot_levels)
ggplot(da_results_sex, aes(ol_clusters_named_fraction)) + geom_histogram(bins=50)
```


```{r}
da_results_sex$clust <- ifelse(da_results_sex$ol_clusters_named_fraction < 0.7, 
                           "Mixed", 
                           da_results_sex$ol_clusters_named_fraction)

sex_beeswarm <- plotDAbeeswarm(da_results_sex, group.by = "ol_clusters_named")
sex_beeswarm
```


```{r}

pdf(file = here(milo_dir_bees, "Sex_corr_for_tissue_age.pdf"),   
              width = 7, 
              height = 13) 
print(sex_beeswarm)
dev.off()
```


# Finding markers of DA populations

## Automatic grouping of neighborhoods

```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Find groups
da_results_sex <- groupNhoods(milo_obj, da_results_sex, max.lfc.delta = 2)
head(da_results_sex)

```

```{r}

plotNhoodGroups(milo_obj, da_results_sex, layout="UMAP") 
```


```{r}

plotDAbeeswarm(da_results_sex, "NhoodGroup")
```


```{r, fig.height = 5, fig.width = 7}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_sex, max.lfc.delta = 0.5), 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=0.5")

```

```{r, fig.height = 5, fig.width = 7}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_sex, max.lfc.delta = 0.2), 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=0.2")

```

```{r}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_sex, max.lfc.delta = 1) , 
               group.by = "NhoodGroup") + ggtitle("max LFC delta=1")


```

```{r}
plotDAbeeswarm(groupNhoods(milo_obj, da_results_sex, max.lfc.delta = 1, 
                           overlap=3), group.by = "NhoodGroup") + 
  ggtitle("overlap=3")

```


```{r}
set.seed(42)
da_results_sex <- groupNhoods(milo_obj, da_results_sex, max.lfc.delta = 1, overlap=3)
plotNhoodGroups(milo_obj, da_results_sex, layout="UMAP")

```



```{r}
plotDAbeeswarm(da_results_sex, group.by = "NhoodGroup")

```
# Finding gene signatures for neighbourhoods

```{r, eval = FALSE}
keep_rows <- rowSums(logcounts(milo_obj)) != 0
milo_obj <- milo_obj[keep_rows, ]

## Find HVGs
dec <- modelGeneVar(milo_obj)
hvgs <- getTopHVGs(dec, n=2000)
head(hvgs)
```



```{r, eval= F}
nhood_markers <- findNhoodGroupMarkers(milo_obj, 
                                       da_results_sex, 
                                       subset.row = hvgs, 
                                       aggregate.samples = TRUE, 
                                       sample_col = "uniq_id")

head(nhood_markers)

save_dir <- here("outs",
                 "Milo_compositional_analysis",
                 "Milo_neighbourhood_mark")
dir.create(save_dir)
write.csv(nhood_markers, here(save_dir,
                              "Milo_neighbour_markers.csv"))
```
# Save data

```{r}
milo_dir <- here("data", "Milo_datasets")
dir.create(milo_dir)
saveRDS(milo_obj, here(milo_dir, "HCA_all_milo.RDS"))

da_data_dir <- here("outs", "Milo_compositional_analysis",
                    "differential_abundance")
dir.create(da_data_dir)

write.csv(da_results, here(da_data_dir, "da_results_tissue.csv"))
write.csv(da_results_age, here(da_data_dir, "da_results_age.csv"))
write.csv(da_results_sex, here(da_data_dir, "da_results_sex.csv"))

```

# Session info
```{r}
sessionInfo()

```

