---
title: "GO_cluster"
author: "Luise A. Seeker"
date: "08/11/2021"
output: html_document
---
# introduction

I tested GSEA and gene ontology both for molecular functions and biological 
pathways and different visualisations. In general GO works better for my data
than GSEA so therefore in later versions of this document, there may be no 
code for GSEA any more. 


```{r}
library(here)
library(tibble)

library(Seurat)
library(dplyr)
library(ggplot2)
library(tidyr)

#GO & GSEA


library(biomaRt)
library(org.Hs.eg.db)
library(ReactomePA)
library(msigdbr)
library(fgsea)
library(NMF)
library(ggupset)
library(clusterProfiler)
library(enrichplot)

```



# Read in data

```{r}

nad_ol <- readRDS(here("data", 
                       "single_nuc_data", 
                       "oligodendroglia",
                       "srt_oligos_and_opcs_LS.RDS"))


Idents(nad_ol) <- "ol_clusters_named" 
opcs <- subset(nad_ol, idents = c("OPC_A", "OPC_B"))

Idents(opcs) <- "Tissue" 

csc_ba4_mark <- FindMarkers(opcs, ident.1 = "CSC",
                          ident.2 = "BA4",
                          only.pos = FALSE, 
                          min.pct = 0.25,
                          test.use = "MAST")
```

# Prepare hallmarks for GSEA and gene sets

```{r, eval = FALSE}


hallmark <- msigdbr(species = "Homo sapiens", category = "H")
geneset <- hallmark %>% split(x = .$gene_symbol, f = .$gs_name)


```





```{r}



ba4 <- subset(csc_ba4_mark, csc_ba4_mark$avg_log2FC < 0)
csc <- subset(csc_ba4_mark, csc_ba4_mark$avg_log2FC >0)


# Prepare marts to convert to entrez
listMarts()
ensembl = useMart("ensembl", dataset="hsapiens_gene_ensembl",
                      host = "www.ensembl.org")
    
    
listDatasets(ensembl)
attributes = listAttributes(ensembl)


 
genes_ba4 <- rownames(ba4)
  
  
# convert to ENTREZID
genes_ent <- bitr(genes_ba4, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
  


# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = genes_ent, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)
```


```{r}


genes_csc <- rownames(csc)
  
  
# convert to ENTREZID
genes_ent <- bitr(genes_csc, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
  


# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = genes_ent, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)



```





```{r}

fil_data <- subset(csc, csc$p_val_adj < 0.05)
    
genes <- rownames(fil_data)

fil_data$abs_avg_log2FC <- abs(fil_data$avg_log2FC)
fil_data$gene <- rownames(fil_data)
  
# Prepare input data
ranks <- fil_data %>% dplyr::select(gene, abs_avg_log2FC)
ranks <- deframe(ranks)

#RUN GO using clusterProfiler
fil_data$Gene <- fil_data$gene
#remove any duplicates (sanity check for me)
genemodulesGO <- fil_data[!duplicated(fil_data$Gene),]
    
Biomart_gencode_ensembl84_biotypes <- getBM(attributes=c("ensembl_gene_id",
                                                             "entrezgene_id",
                                                             "gene_biotype",
                                                             "hgnc_symbol"), 
                                                filters = "hgnc_symbol", 
                                                values = fil_data$Gene, 
                                                mart = ensembl,
                         uniqueRows = FALSE)
    
    
Biomart_gencode_ensembl84_biotypes[, 'gene_biotype'] <- 
      as.factor(Biomart_gencode_ensembl84_biotypes[,'gene_biotype'])
    
    #Filter genes
    #biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
    #                              %in% nad_ol@assays$RNA@var.features)
    #entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
    #                    %in% nad_ol@assays$RNA@var.features)

   # if(nrow(biotype_all_dataset) == 0){
      biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
                                  %in% rownames(opcs@assays$RNA))
      
      entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
                        %in% rownames(opcs@assays$RNA))
   # }
    
    entrezmatched <- entrezID[match(genemodulesGO$Gene,entrezID$hgnc_symbol),]
  
    #you might need to remove NAs
    entrezmatched <- entrezmatched[! apply(entrezmatched[,c(1,3)], 1,function (x) anyNA(x)),]
    allLLIDs <- entrezmatched$entrezgene
    
    modulesReactome <- enrichPathway(gene=as.character(allLLIDs),organism="human",
                                        pvalueCutoff=0.01,
                                        qvalueCutoff = 0.3,
                                        pAdjustMethod = "none", 
                                        readable=T)
    
    
  
    
    if(nrow(as.data.frame(modulesReactome)) > 0){
        dot_plot <- dotplot(modulesReactome, showCategory=20)
        
        pdf(file = here(go_dir, plot_label),   
          width = 12, 
          height = 10) 
        print(dot_plot)
        dev.off()
      
      
        x2 <- pairwise_termsim(modulesReactome) 
        emap_plot <- emapplot(x2)
        
        pdf(file = here(emap_dir, plot_label),   
          width = 12, 
          height = 10) 
        
        print(emap_plot)
        
        dev.off()
        
     #Alternative GO clasissification
        
         ent_uni <- rownames(nad_ol@assays$RNA)
         ent_uni_entrez <- bitr(ent_uni, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
         
         ent_uni <- ent_uni_entrez$ENTREZID
         
         # Molecular function
         
          go_mf <- enrichGO(gene  = as.character(allLLIDs),
                universe      = ent_uni,
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 1,
                qvalueCutoff  = 1,
                readable      = TRUE)
          
          
         
          
          go_mf_plot <- dotplot(go_mf, showCategory = 20)
          pdf(file = here(go_dir_mol_f, plot_label),   
                 width = 12, 
                height = 10) 
          print(go_mf_plot)
          dev.off()
          
          
          
        # GO biological process
        
        ego <- enrichGO(gene  = as.character(allLLIDs),
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                universe = ent_uni )
                
        #head(ego)
        if(nrow(ego) > 1){
        dp2 <- dotplot(ego, showCategory=20)
    
    
        pdf(file = here(go_dot, plot_label),   
            width = 12, 
            height = 10) 
        print(dp2)
        dev.off()
        
        ## convert gene ID to Symbol
        edox <- setReadable(ego, 'org.Hs.eg.db', 'ENTREZID')
        cnet_pl <- cnetplot(edox, foldChange=ranks, circular = TRUE, 
                       colorEdge = TRUE)
    
    
        pdf(file = here(go_dir_cnetpl, plot_label),   
              width = 14, 
              height = 12) 
        print(cnet_pl)
        dev.off()
        
        
        
        edox2 <- enrichplot::pairwise_termsim(edox)
       
        
        if(nrow(edox2) > 1){
        tree_plot <- treeplot(edox2)
        
        pdf(file = here(go_dir_tree, plot_label),   # The directory you want to save the file in
              width = 16, # The width of the plot in inches
              height = 10) # The height of the plot in inches
        print(tree_plot)
        dev.off()
        }
        
       u_plot <- upsetplot(ego)
       
       pdf(file = here(go_dir_upset_plot, plot_label),   
              width = 15, 
              height = 10) 
        print(u_plot)
        dev.off()
       
        if(nrow(edox) > 1){
            hm <- heatplot(edox, foldChange=ranks, showCategory=5)
       
            pdf(file = here(go_dir_hm, plot_label),   # The directory you want to save the file in
              width = 15, # The width of the plot in inches
              height = 10) # The height of the plot in inches
            print(hm)
            dev.off()
        }
        }
     
        
        
        edox_mf <- setReadable(go_mf, 'org.Hs.eg.db', 'ENTREZID')
        edox2_mf <- pairwise_termsim(edox_mf)
        tree_plot_mf <- treeplot(edox2_mf)
        
        pdf(file = here(go_dir_mol_f_tree, plot_label),   
              width = 16, 
              height = 10) 
        print(tree_plot_mf)
        dev.off()
        
    }
  }
}

```


# for non oligos

```{r}
files <- c( here("outs",
                "astrocytes",
                "tissue_marker_list_astro",
                "astro_tissue_marker.csv"),
            
            here("outs",
                "astrocytes",
                "age_marker_list_astro",
                "astro_age_marker.csv"),
           
           here("outs",
                "astrocytes",
                "sex_marker_list_astro",
                "astro_sex_marker.csv"),
           
            here("outs",
                "astrocytes",
                "age_sex_marker_list_astro",
                "astro_age_sex_marker.csv"),
           
           ###
            here("outs",
                "microglia_macrophages",
                "tissue_marker_list_mm",
                "mm_tissue_marker.csv"),
           
            here("outs",
                "microglia_macrophages",
                "age_marker_list_mm",
                "mm_age_marker.csv"),
           
            here("outs",
                "microglia_macrophages",
                "sex_marker_list_mm",
                "mm_sex_marker.csv"),
           
            here("outs",
                "microglia_macrophages",
                "age_sex_marker_list_mm",
                "mm_age_sex_marker.csv"),
           
           ###
           
            here("outs",
                "vascular_cells",
                "tissue_marker_list_vasc",
                "vasc_tissue_mark.csv"),
           
            here("outs",
                "vascular_cells",
                "age_marker_list_vasc",
                "vasc_age_marker.csv"),
           
            here("outs",
                "vascular_cells",
                "sex_marker_list_vasc",
                "vasc_sex_marker.csv"),
           
            here("outs",
                "vascular_cells",
                "age_sex_marker_list_vasc",
                "vasc_age_sex_mark.csv"),
           
           ###
           
            here("outs",
                "neurons",
                "tissue_marker_list_neuron",
                "neuron_tissue_marker.csv"),
           
           here("outs",
                "neurons",
                "age_marker_list_neuron",
                "neuron_age_marker.csv"),
           
           here("outs",
                "neurons",
                "sex_marker_list_neuron",
                "neuron_sex_marker.csv"),
           
           here("outs",
                "neurons",
                "age_sex_marker_list_neuron",
                "age_sex_neurons_marker.csv")
           
           )


single_nuc_data <- c(here("data",
                          "single_nuc_data",
                          "astrocytes",
                          "HCA_astrocytes.RDS"),
                     
                     here("data",
                          "single_nuc_data",
                          "microglia",
                          "HCA_microglia.RDS"),
                     
                     here("data",
                          "single_nuc_data",
                          "vascular_cells",
                          "HCA_vascular_cells.RDS"),
                     
                     here("data",
                          "single_nuc_data",
                          "neurons",
                          "HCA_neurons.RDS")
  
  
)
                     
```

```{r}

named_list <- list()

for(i in 1: length(files)){
  
  data <- read.csv(files[i])
  cluster_lev <- levels(as.factor(data$cluster))
  for(k in 1:length(cluster_lev)){
    curr_data <- subset(data, data$cluster == cluster_lev[k])
    cluster <- curr_data[["cluster"]][1]
    cell_t <- sapply(strsplit(files[i], "/"), `[`, 12)
    cell_t <- sapply(strsplit(cell_t, "_marker"), `[`, 1)
    label <- paste0(cell_t, "_", cluster)
    plot_label <- paste0(label, ".pdf")
    fil_data <- subset(curr_data, curr_data$p_val_adj < 0.05 &
                       curr_data$avg_log2FC > 0)
    
    genes <- fil_data$gene
  
  
  # convert to ENTREZID
  
    genes_ent <- bitr(genes, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
  
    named_list[[k]] <- genes_ent$ENTREZID
    names(named_list)[k] <- label
  }
  if(i == 1){
    return_list <- named_list
  }else{
    return_list <- append(return_list, named_list)
  }
  named_list <- list()
}

str(return_list)

tissue_list_astro <- return_list[1:3]
age_list_astro <- return_list[4:5]
sex_list_astro <- return_list[6:7]
age_sex_list_astro <- return_list[8:11]

tissue_list_mm <- return_list[12:14]
age_list_mm <- return_list[15:16]
sex_list_mm <- return_list[17:18]
age_sex_list_mm <- return_list[19:22]


tissue_list_vasc <- return_list[23:25]
age_list_vasc <- return_list[26:27]
sex_list_vasc <- return_list[28:29]
age_sex_list_vasc <- return_list[30:33]

tissue_list_neur <- return_list[34:36]
age_list_neur <- return_list[37:38]
sex_list_neur <- return_list[39:40]
age_sex_list_neur <- return_list[41:44]




# ASTROCYTES
# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = tissue_list_astro, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)

# Age
ck_kegg_age <- compareCluster(geneCluster = age_list_astro, fun = enrichKEGG)
ck_kegg_age <- setReadable(ck_kegg_age, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age) 

dotplot(ck_kegg_age)

# Age*sex
ck_kegg_age_sex <- compareCluster(geneCluster = age_sex_list_astro, fun = enrichKEGG)
ck_kegg_age_sex <- setReadable(ck_kegg_age_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age_sex) 

dotplot(ck_kegg_age_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# sex
ck_kegg_sex <- compareCluster(geneCluster = sex_list_astro, fun = enrichKEGG)
ck_kegg_sex <- setReadable(ck_kegg_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_sex) 

dotplot(ck_kegg_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



# MICROGLIA
# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = tissue_list_mm, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Age
ck_kegg_age <- compareCluster(geneCluster = age_list_mm, fun = enrichKEGG)
ck_kegg_age <- setReadable(ck_kegg_age, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age) 

dotplot(ck_kegg_age)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Age*sex
ck_kegg_age_sex <- compareCluster(geneCluster = age_sex_list_mm, fun = enrichKEGG)
ck_kegg_age_sex <- setReadable(ck_kegg_age_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age_sex) 

dotplot(ck_kegg_age_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# sex
ck_kegg_sex <- compareCluster(geneCluster = sex_list_mm, fun = enrichKEGG)
ck_kegg_sex <- setReadable(ck_kegg_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_sex) 

dotplot(ck_kegg_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))




# VASCULAR
# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = tissue_list_vasc, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Age
ck_kegg_age <- compareCluster(geneCluster = age_list_vasc, fun = enrichKEGG)
ck_kegg_age <- setReadable(ck_kegg_age, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age) 

dotplot(ck_kegg_age)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Age*sex
ck_kegg_age_sex <- compareCluster(geneCluster = age_sex_list_vasc, fun = enrichKEGG)
ck_kegg_age_sex <- setReadable(ck_kegg_age_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age_sex) 

dotplot(ck_kegg_age_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# sex
ck_kegg_sex <- compareCluster(geneCluster = sex_list_vasc, fun = enrichKEGG)
ck_kegg_sex <- setReadable(ck_kegg_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_sex) 

dotplot(ck_kegg_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



# neurons
# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = tissue_list_neur, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Age
ck_kegg_age <- compareCluster(geneCluster = age_list_neur, fun = enrichKEGG)
ck_kegg_age <- setReadable(ck_kegg_age, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age) 

dotplot(ck_kegg_age)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Age*sex
ck_kegg_age_sex <- compareCluster(geneCluster = age_sex_list_neur, fun = enrichKEGG)
ck_kegg_age_sex <- setReadable(ck_kegg_age_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_age_sex) 

dotplot(ck_kegg_age_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# sex
ck_kegg_sex <- compareCluster(geneCluster = sex_list_neur, fun = enrichKEGG)
ck_kegg_sex <- setReadable(ck_kegg_sex, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_sex) 

dotplot(ck_kegg_sex) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```





```{r}

for(i in 1: length(files)){
  data <- read.csv(files[i])
  cluster_lev <- levels(as.factor(data$cluster))
  for(k in 1:length(cluster_lev)){
    curr_data <- subset(data, data$cluster == cluster_lev[k])
    cluster <- curr_data[["cluster"]][1]
    cluster <- gsub(" ", "_", cluster)
    cell_t <- sapply(strsplit(files[i], "/"), `[`, 12)
    cell_t <- sapply(strsplit(cell_t, "_mark"), `[`, 1)
    label <- paste0(cell_t, "_", cluster)
    plot_label <- paste0(label, ".pdf")
    cell_type <- sapply(strsplit(cell_t, "_"), `[`, 1)
    fil_data <- subset(curr_data, curr_data$p_val_adj < 0.05 &
                       curr_data$avg_log2FC > 0)
    if(nrow(fil_data) > 0){
      
    
    genes <- fil_data$gene
  
  
    # Prepare input data
    ranks <- fil_data %>% dplyr::select(gene, avg_log2FC)
    ranks <- deframe(ranks)

  #RUN GO using clusterProfiler
    fil_data$Gene <- fil_data$gene
    #remove any duplicates (sanity check for me)
    genemodulesGO <- fil_data[!duplicated(fil_data$Gene),]
    
    Biomart_gencode_ensembl84_biotypes <- getBM(attributes=c("ensembl_gene_id",
                                                             "entrezgene_id",
                                                             "gene_biotype",
                                                             "hgnc_symbol"), 
                                                filters = "hgnc_symbol", 
                                                values = fil_data$Gene, 
                                                mart = ensembl,
                         uniqueRows = FALSE)
    
    
    Biomart_gencode_ensembl84_biotypes[, 'gene_biotype'] <- 
      as.factor(Biomart_gencode_ensembl84_biotypes[,'gene_biotype'])
    
    
    if(cell_type == "astro"){
      srt <- readRDS(single_nuc_data[1])
    }else if(cell_type == "mm"){
      srt <- readRDS(single_nuc_data[2])
    }else if(cell_type == "vasc"){
      srt <- readRDS(single_nuc_data[3])
    }else if(cell_type == "neuron" | cell_type == "age"){
      srt <- readRDS(single_nuc_data[4])
    }else{
      print("check srt.")
    }
    
    #Filter genes
    #biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
    #                              %in% srt@assays$RNA@var.features)
    #entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
    #                    %in% srt@assays$RNA@var.features)

   # if(nrow(biotype_all_dataset) == 0){
      biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
                                  %in% rownames(srt@assays$RNA))
      
      entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
                        %in% rownames(srt@assays$RNA))
   # }
    
    entrezmatched <- entrezID[match(genemodulesGO$Gene,entrezID$hgnc_symbol),]
  
    #you might need to remove NAs
    entrezmatched <- entrezmatched[! apply(entrezmatched[,c(1,3)], 1,function (x) anyNA(x)),]
    allLLIDs <- entrezmatched$entrezgene
    
    modulesReactome <- enrichPathway(gene=as.character(allLLIDs),organism="human",
                                        pvalueCutoff=0.01,
                                        qvalueCutoff = 0.3,
                                        pAdjustMethod = "none", 
                                        readable=T)
    
    
  
    
    if(nrow(as.data.frame(modulesReactome)) > 0){
        dot_plot <- dotplot(modulesReactome, showCategory=20)
        
        pdf(file = here(go_dir, plot_label),   
          width = 12, 
          height = 10) 
        print(dot_plot)
        dev.off()
      
      
        x2 <- pairwise_termsim(modulesReactome) 
        emap_plot <- emapplot(x2)
        
        pdf(file = here(emap_dir, plot_label),   
          width = 12, 
          height = 10) 
        
        print(emap_plot)
        
        dev.off()
        
     #Alternative GO clasissification
        
         ent_uni <- rownames(srt@assays$RNA)
         ent_uni_entrez <- bitr(ent_uni, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
         
         ent_uni <- ent_uni_entrez$ENTREZID
         
         # Molecular function
         
          go_mf <- enrichGO(gene  = as.character(allLLIDs),
                universe      = ent_uni,
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 1,
                qvalueCutoff  = 1,
                readable      = TRUE)
         
          
          go_mf_plot <- dotplot(go_mf, showCategory = 20)
          pdf(file = here(go_dir_mol_f, plot_label),   
                 width = 12, 
                height = 10) 
          print(go_mf_plot)
          dev.off()
          
        # GO biological process
        
        ego <- enrichGO(gene  = as.character(allLLIDs),
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                universe = ent_uni )
                
        #head(ego)
        if(nrow(ego) > 1){
        dp2 <- dotplot(ego, showCategory=20)
    
    
        pdf(file = here(go_dot, plot_label),   
            width = 12, 
            height = 10) 
        print(dp2)
        dev.off()
        
        ## convert gene ID to Symbol
        edox <- setReadable(ego, 'org.Hs.eg.db', 'ENTREZID')
        cnet_pl <- cnetplot(edox, foldChange=ranks, circular = TRUE, 
                       colorEdge = TRUE)
    
    
        pdf(file = here(go_dir_cnetpl, plot_label),   
              width = 14, 
              height = 12) 
        print(cnet_pl)
        dev.off()
        
        
        
       # edox2 <- enrichplot::pairwise_termsim(edox)
       
        
       # if(nrow(edox2) > 1){
       # tree_plot <- treeplot(edox2)
        
        #pdf(file = here(go_dir_tree, plot_label),   # The directory you want to save the file in
       #       width = 16, # The width of the plot in inches
        #      height = 10) # The height of the plot in inches
       # print(tree_plot)
        #dev.off()
        
        
       u_plot <- upsetplot(ego)
       
       pdf(file = here(go_dir_upset_plot, plot_label),   
              width = 15, 
              height = 10) 
        print(u_plot)
        dev.off()
        
        }
       
        if(nrow(edox) > 1){
            hm <- heatplot(edox, foldChange=ranks, showCategory=5)
       
            pdf(file = here(go_dir_hm, plot_label),   # The directory you want to save the file in
              width = 15, # The width of the plot in inches
              height = 10) # The height of the plot in inches
            print(hm)
            dev.off()
        }
        }
     
        
        
        edox_mf <- setReadable(go_mf, 'org.Hs.eg.db', 'ENTREZID')
        edox2_mf <- pairwise_termsim(edox_mf)
        tree_plot_mf <- treeplot(edox2_mf)
        
        pdf(file = here(go_dir_mol_f_tree, plot_label),   
              width = 16, 
              height = 10) 
        print(tree_plot_mf)
        dev.off()
        
    }
  }
}


```



# Create and save files that contain GO output for oligos

```{r}



oligo_files
srt <- nad_ol

plot_df <- data.frame()
for(i in 1: length(oligo_files)){
  data <- read.csv(oligo_files[i])
  #srt <- readRDS(single_nuc_data[i])
  cluster_lev <- levels(as.factor(data$cluster))
  for(k in 1:length(cluster_lev)){
    curr_data <- subset(data, data$cluster == cluster_lev[k])
    cluster <- curr_data[["cluster"]][1]
    #cell_t <- sapply(strsplit(oligo_files[i], "/"), `[`, 10)
    cell_t <- "oligos_"
    plot_label <- paste0(cell_t, "cluster", "GO_term_age_sex_region.csv")
    fil_data <- subset(curr_data, curr_data$p_val_adj < 0.05 &
                       curr_data$avg_log2FC > 0)

   #RUN GO using clusterProfiler
    fil_data$Gene <- fil_data$gene
    #remove any duplicates (sanity check for me)
    genemodulesGO <- fil_data[!duplicated(fil_data$Gene),]
    
    Biomart_gencode_ensembl84_biotypes <- getBM(attributes=c("ensembl_gene_id",
                                                             "entrezgene_id",
                                                             "gene_biotype",
                                                             "hgnc_symbol"), 
                                                filters = "hgnc_symbol", 
                                                values = fil_data$Gene, 
                                                mart = ensembl,
                         uniqueRows = FALSE)
    
    
    Biomart_gencode_ensembl84_biotypes[, 'gene_biotype'] <- 
      as.factor(Biomart_gencode_ensembl84_biotypes[,'gene_biotype'])
    

    biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
                                  %in% rownames(srt@assays$RNA))
      
    entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
                        %in% rownames(srt@assays$RNA))
  
    
    entrezmatched <- entrezID[match(genemodulesGO$Gene,entrezID$hgnc_symbol),]
  
    #you might need to remove NAs
    entrezmatched <- entrezmatched[! apply(entrezmatched[,c(1,3)], 1,function (x) anyNA(x)),]
    allLLIDs <- entrezmatched$entrezgene
    allLLIDs<-allLLIDs[!is.na(allLLIDs)]
    ent_uni <- rownames(nad_ol@assays$RNA)
    ent_uni_entrez <- bitr(ent_uni, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
         
    ent_uni <- ent_uni_entrez$ENTREZID
         
         
        
   ego <- enrichGO(gene  = as.character(allLLIDs),
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                universe = ent_uni )
   
   ego_df <- as.data.frame(ego)
   
   if(nrow(ego_df)> 0){
    ego_df$Cell_type <- cell_t
    ego_df$cluster <- cluster_lev[k]
   
    if(nrow(plot_df) < 1){
      plot_df <- ego_df
    }else{
      plot_df <- rbind(plot_df, ego_df)
    }
   }
   
  }

}


dir.create(here("outs",
                       "GO",
                       "tissue_age_sex_table_output",
                       "oligos"), recursive = TRUE)

write.csv(plot_df, here("outs",
                       "GO",
                       "tissue_age_sex_table_output",
                       "oligos",
                       paste0(plot_label)))


                

```


# Create and save files that contain GO output for non_oligos

```{r}


 

plot_df <- data.frame()


for(i in 1: length(files)){
  data <- read.csv(files[i])
  cell_t <- sapply(strsplit(files[i], "/"), `[`, 10)
  
  if(cell_t == "astrocytes"){
      srt <- readRDS(single_nuc_data[1])
    }else if(cell_t == "microglia_macrophages"){
      srt <- readRDS(single_nuc_data[2])
    }else if(cell_t == "vascular_cells"){
      srt <- readRDS(single_nuc_data[3])
    }else if(cell_t == "neurons"){
      srt <- readRDS(single_nuc_data[4])
    }else{
      print("check srt.")
    }
  
  cluster_lev <- levels(as.factor(data$cluster))
  
  for(k in 1:length(cluster_lev)){
    curr_data <- subset(data, data$cluster == cluster_lev[k])
    cluster <- curr_data[["cluster"]][1]
    plot_label <- paste0(cell_t, "_", cluster, "_GO_term_age_sex_region.csv")
    fil_data <- subset(curr_data, curr_data$p_val_adj < 0.05 &
                       curr_data$avg_log2FC > 0)
    
    if(nrow(fil_data) > 0){

   #RUN GO using clusterProfiler
    fil_data$Gene <- fil_data$gene
    #remove any duplicates (sanity check for me)
    genemodulesGO <- fil_data[!duplicated(fil_data$Gene),]
    
    Biomart_gencode_ensembl84_biotypes <- getBM(attributes=c("ensembl_gene_id",
                                                             "entrezgene_id",
                                                             "gene_biotype",
                                                             "hgnc_symbol"), 
                                                filters = "hgnc_symbol", 
                                                values = fil_data$Gene, 
                                                mart = ensembl,
                         uniqueRows = FALSE)
    
    
    Biomart_gencode_ensembl84_biotypes[, 'gene_biotype'] <- 
      as.factor(Biomart_gencode_ensembl84_biotypes[,'gene_biotype'])
    

    biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
                                  %in% rownames(srt@assays$RNA))
      
    entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
                        %in% rownames(srt@assays$RNA))
  
    
    entrezmatched <- entrezID[match(genemodulesGO$Gene,entrezID$hgnc_symbol),]
  
    #you might need to remove NAs
    entrezmatched <- entrezmatched[! apply(entrezmatched[,c(1,3)], 1,function (x) anyNA(x)),]
    allLLIDs <- entrezmatched$entrezgene
    allLLIDs<-allLLIDs[!is.na(allLLIDs)]
    ent_uni <- rownames(nad_ol@assays$RNA)
    ent_uni_entrez <- bitr(ent_uni, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
         
    ent_uni <- ent_uni_entrez$ENTREZID
         
         
        
   ego <- enrichGO(gene  = as.character(allLLIDs),
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                universe = ent_uni )
   
   ego_df <- as.data.frame(ego)
   
   if(nrow(ego_df)> 0){
    ego_df$Cell_type <- cell_t
    ego_df$cluster <- cluster_lev[k]
   
    if(nrow(plot_df) < 1){
      plot_df <- ego_df
    }else{
      plot_df <- rbind(plot_df, ego_df)
    }
   }
   }
  }
}


dir.create(here("outs",
                       "GO",
                       "tissue_age_sex_table_output",
                       "non_oligos"), recursive = TRUE)

write.csv(plot_df, here("outs",
                       "GO",
                       "tissue_age_sex_table_output",
                       "non_oligos",
                       "non_oligos_GO_term_age_sex_region.csv"))


                

```





# Session Info
```{r}
sessionInfo()

```


