---
title: "GO_cluster"
author: "Luise A. Seeker"
date: "08/11/2021"
output: html_document
---
# introduction

I tested GSEA and gene ontology both for molecular functions and biological 
pathways and different visualisations. In general GO works better for my data
than GSEA so therefore in later versions of this document, there may be no 
code for GSEA any more. 


```{r}
library(here)
library(tibble)

library(Seurat)
library(dplyr)
library(ggplot2)
library(tidyr)
library(EnhancedVolcano)
library(ggsci)

#GO & GSEA


library(biomaRt)
library(org.Hs.eg.db)
library(ReactomePA)
library(msigdbr)
library(fgsea)
library(NMF)
library(ggupset)
library(clusterProfiler)
library(enrichplot)
library(stringr)

```

```{r, echo = F}
mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)

mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)

```

# Read in data

```{r}

nad_ol <- readRDS(here("data", 
                       "single_nuc_data", 
                       "oligodendroglia",
                       "srt_oligos_and_opcs_LS.RDS"))


Idents(nad_ol) <- "ol_clusters_named" 
opcs <- subset(nad_ol, idents = c("OPC_A", "OPC_B"))

Idents(opcs) <- "Tissue" 

csc_ba4_mark <- FindMarkers(opcs, ident.1 = "CSC",
                          ident.2 = "BA4",
                          only.pos = FALSE, 
                          min.pct = 0.2,
                          test.use = "MAST")
```
```{r, fig.height = 4, fig.width = 4}

bool_hox<- str_detect(rownames(csc_ba4_mark), pattern = "^HOX", negate = FALSE)

EnhancedVolcano(csc_ba4_mark,
    lab = rownames(csc_ba4_mark),
    x = 'avg_log2FC',
    y = 'p_val_adj',
    title = 'CSC OPCs vs. BA4 OPCs',
    pCutoff = 0.05,
    FCcutoff = 0.2,
    pointSize = 3.0,
    labSize = 6.0,
    col=c(mycoloursP),
    colAlpha = 4/5,
    drawConnectors = TRUE,
    widthConnectors = 1.0,
    colConnectors = 'black',
    selectLab = c("FOXG1","HOXB3", "NELL1",
                  "L3MBTL4", "LRRC7",
                  "TMEM196", 
                  "DACH2",
                  "MYT1L", "CACNG5",
                  "SEMA3D", "PAX3", 
                  "CDH8",
                  "EBF1", 
                  "HS3ST1", "PDZRN3"))

```


# Prepare hallmarks for GSEA and gene sets

```{r, eval = FALSE}


hallmark <- msigdbr(species = "Homo sapiens", category = "H")
geneset <- hallmark %>% split(x = .$gene_symbol, f = .$gs_name)


```

# 1) KEGG

```{r}



ba4 <- subset(csc_ba4_mark, csc_ba4_mark$avg_log2FC < 0)
csc <- subset(csc_ba4_mark, csc_ba4_mark$avg_log2FC >0)


# Prepare marts to convert to entrez
listMarts()
ensembl = useMart("ensembl", dataset="hsapiens_gene_ensembl",
                      host = "www.ensembl.org")
    
    
listDatasets(ensembl)
attributes = listAttributes(ensembl)

```


BA4 vs CSC OPCs
```{r}
 
genes_ba4 <- rownames(ba4)
  
  
# convert to ENTREZID
genes_ent <- bitr(genes_ba4, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
  


# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = genes_ent, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)
```

CSC vs BA4 OPCs
```{r}


genes_csc <- rownames(csc)
  
  
# convert to ENTREZID
genes_ent <- bitr(genes_csc, fromType = "SYMBOL",
            toType = "ENTREZID",
            OrgDb = org.Hs.eg.db)
  


# Tissue
ck_kegg_tissue <- compareCluster(geneCluster = genes_ent, fun = enrichKEGG)
ck_kegg_tissue <- setReadable(ck_kegg_tissue, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck_kegg_tissue) 

dotplot(ck_kegg_tissue)



```


# 2) Cell Profiler

BA4 vs. CSC

```{r}

fil_data_ba4 <- subset(ba4, ba4$p_val_adj < 0.05)
    
genes <- rownames(fil_data_ba4)

fil_data_ba4$abs_avg_log2FC <- abs(fil_data_ba4$avg_log2FC)
fil_data_ba4$gene <- rownames(fil_data_ba4)
  
# Prepare input data
ranks <- fil_data_ba4 %>% dplyr::select(gene, abs_avg_log2FC)
ranks <- deframe(ranks)

#RUN GO using clusterProfiler
fil_data_ba4$Gene <- fil_data_ba4$gene
#remove any duplicates (sanity check for me)
genemodulesGO <- fil_data_ba4[!duplicated(fil_data_ba4$Gene),]
    
Biomart_gencode_ensembl84_biotypes <- getBM(attributes=c("ensembl_gene_id",
                                                             "entrezgene_id",
                                                             "gene_biotype",
                                                             "hgnc_symbol"), 
                                                filters = "hgnc_symbol", 
                                                values = fil_data_ba4$Gene, 
                                                mart = ensembl,
                         uniqueRows = FALSE)
    
    
Biomart_gencode_ensembl84_biotypes[, 'gene_biotype'] <- 
      as.factor(Biomart_gencode_ensembl84_biotypes[,'gene_biotype'])
    

biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
                                  %in% rownames(opcs@assays$RNA))
      
entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
                        %in% rownames(opcs@assays$RNA))
   
entrezmatched <- entrezID[match(genemodulesGO$Gene,entrezID$hgnc_symbol),]
  
    #you might need to remove NAs
entrezmatched <- entrezmatched[! apply(entrezmatched[,c(1,3)], 1,function (x) anyNA(x)),]
allLLIDs <- entrezmatched$entrezgene
    
modulesReactome <- enrichPathway(gene=as.character(allLLIDs),organism="human",
                                        pvalueCutoff=0.01,
                                        qvalueCutoff = 0.3,
                                        pAdjustMethod = "none", 
                                        readable=T)
    
    
  
    

dotplot(modulesReactome, showCategory=20)

```


```{r}
x2 <- pairwise_termsim(modulesReactome) 
emapplot(x2)
        
```

```{r}

ent_uni <- rownames(opcs@assays$RNA)
ent_uni_entrez <- bitr(ent_uni, fromType = "SYMBOL",
                       toType = "ENTREZID",
                       OrgDb = org.Hs.eg.db)
         
ent_uni <- ent_uni_entrez$ENTREZID
         
# Molecular function
         
go_mf <- enrichGO(gene  = as.character(allLLIDs),
                universe      = ent_uni,
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 1,
                qvalueCutoff  = 1,
                readable      = TRUE)
          
          
         
          
dotplot(go_mf, showCategory = 15)

```


```{r}
ego <- enrichGO(gene  = as.character(allLLIDs),
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                universe = ent_uni )
#dotplot(ego, showCategory=20)
```

ego has nrow of 0


```{r}

## convert gene ID to Symbol
#edox <- setReadable(ego, 'org.Hs.eg.db', 'ENTREZID')
#cnetplot(edox, foldChange=ranks, circular = TRUE, 
#                       colorEdge = TRUE)
```


```{r}
#edox2 <- enrichplot::pairwise_termsim(edox)
#treeplot(edox2)
```

```{r}
#u_plot <- upsetplot(ego)
```

```{r}
#heatplot(edox, foldChange=ranks, showCategory=5)
```

```{r}
edox_mf <- setReadable(go_mf, 'org.Hs.eg.db', 'ENTREZID')
edox2_mf <- pairwise_termsim(edox_mf)
treeplot(edox2_mf)
  
```

```{r}
heatplot(edox2_mf, foldChange=ranks, showCategory=5)
```


CSC vs NA4 OPCs
```{r}

fil_data <- subset(csc, csc$p_val_adj < 0.05)
    
genes <- rownames(fil_data)

fil_data$abs_avg_log2FC <- abs(fil_data$avg_log2FC)
fil_data$gene <- rownames(fil_data)
  
# Prepare input data
ranks <- fil_data %>% dplyr::select(gene, abs_avg_log2FC)
ranks <- deframe(ranks)

#RUN GO using clusterProfiler
fil_data$Gene <- fil_data$gene
#remove any duplicates (sanity check for me)
genemodulesGO <- fil_data[!duplicated(fil_data$Gene),]
    
Biomart_gencode_ensembl84_biotypes <- getBM(attributes=c("ensembl_gene_id",
                                                             "entrezgene_id",
                                                             "gene_biotype",
                                                             "hgnc_symbol"), 
                                                filters = "hgnc_symbol", 
                                                values = fil_data$Gene, 
                                                mart = ensembl,
                         uniqueRows = FALSE)
    
    
Biomart_gencode_ensembl84_biotypes[, 'gene_biotype'] <- 
      as.factor(Biomart_gencode_ensembl84_biotypes[,'gene_biotype'])
    
    #Filter genes
    #biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
    #                              %in% nad_ol@assays$RNA@var.features)
    #entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
    #                    %in% nad_ol@assays$RNA@var.features)

   # if(nrow(biotype_all_dataset) == 0){
biotype_all_dataset <- subset(Biomart_gencode_ensembl84_biotypes, hgnc_symbol 
                                  %in% rownames(opcs@assays$RNA))
      
entrezID <-  subset(biotype_all_dataset, biotype_all_dataset$hgnc_symbol
                        %in% rownames(opcs@assays$RNA))
   # }
    
entrezmatched <- entrezID[match(genemodulesGO$Gene,entrezID$hgnc_symbol),]
  
    #you might need to remove NAs
entrezmatched <- entrezmatched[! apply(entrezmatched[,c(1,3)], 1,function (x) anyNA(x)),]
allLLIDs <- entrezmatched$entrezgene
    
modulesReactome <- enrichPathway(gene=as.character(allLLIDs),organism="human",
                                        pvalueCutoff=0.01,
                                        qvalueCutoff = 0.3,
                                        pAdjustMethod = "none", 
                                        readable=T)
    
    
  
```

```{r, fig.height =5}

dotplot(modulesReactome, showCategory=20)

```


```{r}
x2 <- pairwise_termsim(modulesReactome) 
emapplot(x2)
        
```

```{r}

ent_uni <- rownames(opcs@assays$RNA)
ent_uni_entrez <- bitr(ent_uni, fromType = "SYMBOL",
                       toType = "ENTREZID",
                       OrgDb = org.Hs.eg.db)
         
ent_uni <- ent_uni_entrez$ENTREZID
         
# Molecular function
         
go_mf <- enrichGO(gene  = as.character(allLLIDs),
                universe      = ent_uni,
                OrgDb         = org.Hs.eg.db,
                ont           = "MF",
                pAdjustMethod = "BH",
                pvalueCutoff  = 1,
                qvalueCutoff  = 1,
                readable      = TRUE)
          
          
         
          
dotplot(go_mf, showCategory = 15)

```

Biological process
```{r}
ego <- enrichGO(gene  = as.character(allLLIDs),
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                universe = ent_uni )
dotplot(ego, showCategory=20)
```


```{r}

## convert gene ID to Symbol
edox <- setReadable(ego, 'org.Hs.eg.db', 'ENTREZID')
cnetplot(edox, foldChange=ranks, circular = TRUE, 
                       colorEdge = TRUE)
```


```{r}
edox2 <- enrichplot::pairwise_termsim(edox)
enrichplot::treeplot(edox2, showCategory = 25,
                     nWords = 10,
                     nCluster = 6,
                     offset = 16.5,
                     offset_tiplab = 0.5,
                     group_color = c("cadetblue",
                                     "cyan4",
                                     "darkgoldenrod",
                                     "cornflowerblue",
                                     "aquamarine4",
                                     "azure4"))
```

```{r}
u_plot <- upsetplot(ego)
```

```{r}
heatplot(edox, foldChange=ranks, showCategory=5)
```

```{r}
edox_mf <- setReadable(go_mf, 'org.Hs.eg.db', 'ENTREZID')
edox2_mf <- pairwise_termsim(edox_mf)
treeplot(edox2_mf)
  
```


```{r}
edox2_df <- edox2@result
edox2_df_fil <- subset(edox2_df, edox2_df$p.adjust < 0.05)
```







```{r}
opcs[["percent.rb"]] <- PercentageFeatureSet(opcs, pattern = "^RP[SL]")


VlnPlot(opcs, 
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "total_percent_mito",
                     "percent.rb"),
        ncol = 4,pt.size = 0.1) & 
  theme(plot.title = element_text(size=10))

```



# Session Info
```{r}
sessionInfo()

```


