---
title: "Validation_gpnmb_p1621_iba1_OLIG2_IF"
author: "Luise A. Seeker"
date: "09/08/2022"
output: html_document
---
HCN2 SPARC and OLIG2 co-labelling in 3 BA4 and 2 CSC samples


```{r}

library(ggplot2)
library(lme4)
library(lmerTest)
library(here)
library(tidyr)
library(ggsci)
library(Seurat)
```

```{r, echo = F}
mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)

mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)

```




```{r}
gpnmb_p1621_iba1 <- read.csv(here("data", 
                   "validation_data", 
                   "20220820_GPNMB_P16_P21_IBA1_FOV_thresholds_quant.csv"))

```


```{r}
names(gpnmb_p1621_iba1)

gpnmb_p1621_iba1$tissue <- as.factor(gpnmb_p1621_iba1$tissue)


gpnmb_p1621_iba1$sample_id <- factor(gpnmb_p1621_iba1$sample_id,
                                     levels = c("SD008_18_BA4",
                                                "SD021_17_BA4",
                                                "SD042_18_BA4",
                                                "SD015_12_CB",
                                                "SD030_12_CB",
                                                "SD039_14_CB",
                                                "SD015_12_CSC",
                                                "SD042_10_CSC"))
```

Convert from wide to long format to see how many single, double and triple
positive cells are in each sample

```{r}

keycol <- "sample_id"
valuecol <- "count"
gathercols <- c("Num_Detections",
                "Num_GPNMB",
                "Num_GPNMB_IBA1", 
                "Num_GPNMB_IBA1_P16P21",
                "Num_GPNMB_P16P21",
                "Num_IBA1",
                "Num_P16P21",
                "Num_P16P21_IBA1")

long_data <- gather(gpnmb_p1621_iba1, keycol, valuecol, gathercols)

head(long_data)
nrow(long_data)
names(long_data)
```






```{r}


ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="stack", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_fill_manual(values= mycoloursP[25:50])





ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="fill", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +scale_fill_manual(values= mycoloursP[25:50])


```
Remove cells that are negative for all tested markers:


```{r}

keycol <- "sample_id"
valuecol <- "count"
gathercols <- c("Num_GPNMB",
                "Num_GPNMB_IBA1", 
                "Num_GPNMB_IBA1_P16P21",
                "Num_GPNMB_P16P21",
                "Num_IBA1",
                "Num_P16P21",
                "Num_P16P21_IBA1")

long_data <- gather(gpnmb_p1621_iba1, keycol, valuecol, gathercols)

head(long_data)
nrow(long_data)
names(long_data)


ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="stack", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_fill_manual(values= mycoloursP[25:50])





ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="fill", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +scale_fill_manual(values= mycoloursP[25:50])

```


Exclude cells that are not positive for OLIG2
```{r}

keycol <- "sample_id"
valuecol <- "count"
gathercols <- c("Num_GPNMB_IBA1", 
                "Num_GPNMB_IBA1_P16P21",
                "Num_IBA1",
                "Num_P16P21_IBA1")

long_data <- gather(gpnmb_p1621_iba1, keycol, valuecol, gathercols)

ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="stack", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_fill_manual(values= mycoloursP[25:50])





ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="fill", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +scale_fill_manual(values= mycoloursP[25:50])


```
collapse tissue
```{r}

ggplot(long_data, aes(x = tissue , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="stack", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_fill_manual(values= mycoloursP[25:50])





ggplot(long_data, aes(x = tissue, 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="fill", stat="identity")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values= mycoloursP[25:50])


```

```{r}

keycol <- "sample_id"
valuecol <- "count"
gathercols <- c("sum_GPNMB_neg_IBA1_Pos",
                "sum_GPNMB_IBA1")

long_data <- gather(gpnmb_p1621_iba1, keycol, valuecol, gathercols)

long_data$keycol <- factor(long_data$keycol, levels = c("sum_GPNMB_neg_IBA1_Pos",
                                                        "sum_GPNMB_IBA1"))

ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="stack", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_fill_manual(values= mycoloursP[25:50])





ggplot(long_data, aes(x = sample_id , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="fill", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +scale_fill_manual(values= mycoloursP[25:50])


```
```{r}

ggplot(long_data, aes(x = tissue , 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="stack", stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+scale_fill_manual(values= mycoloursP[25:50])





ggplot(long_data, aes(x = tissue, 
                          y = valuecol, 
                          fill = keycol)) +
  geom_bar(position="fill", stat="identity")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values= mycoloursP[25:50])


```
```{r}
ggplot(gpnmb_p1621_iba1, aes(x = tissue, y = GPNMB_perc_IBA1)) +
  geom_boxplot()+ geom_jitter(width = 0.2, aes(colour = sample_id)) +
  theme_bw(19) + ylab("GPNMB+IBA1+HOECHST+ (%)") + xlab("Tissue")

```




```{r}

lm_prot <- lm(gpnmb_p1621_iba1$GPNMB_perc_IBA1~ gpnmb_p1621_iba1$tissue)
summary(lm_prot)
anova(lm_prot)
```

```{r}

ti_csc <- subset(gpnmb_p1621_iba1, gpnmb_p1621_iba1$tissue == "CSC")
ti_cb <- subset(gpnmb_p1621_iba1, gpnmb_p1621_iba1$tissue == "CB")
ti_ba4 <- subset(gpnmb_p1621_iba1, gpnmb_p1621_iba1$tissue == "BA4")

t.test(ti_csc$GPNMB_perc_IBA1, ti_cb$GPNMB_perc_IBA1)
t.test(ti_csc$GPNMB_perc_IBA1, ti_ba4$GPNMB_perc_IBA1)
t.test(ti_cb$GPNMB_perc_IBA1, ti_ba4$GPNMB_perc_IBA1)
```


```{r}
summary(gpnmb_p1621_iba1$GPNMB_perc_IBA1)
```

```{r}
micro_srt <-  readRDS(here("data", 
                           "single_nuc_data", 
                           "microglia",
                           "HCA_microglia.RDS"))

Idents(micro_srt) <- "microglia_clu"

micro_srt <- subset(micro_srt, idents = c("Microglia_1",
                                    "Microglia_2",
                                    "Microglia_3",
                                    "Microglia_4",
                                    "Microglia_5",
                                    "BAM"))



Idents(micro_srt) <- "Tissue"

csc_srt <- subset(micro_srt, idents = "CSC")
cb_srt <- subset(micro_srt, idents = "CB")
ba4_srt <- subset(micro_srt, idents = "BA4")

cd_genes <- c("GPNMB")


a_csc <- DotPlot(object = csc_srt, features = cd_genes)
a_csc$data
mean(a_csc$data$pct.exp)


a_cb <- DotPlot(object = cb_srt, features = cd_genes)
a_cb$data
mean(a_cb$data$pct.exp)

a_ba4 <- DotPlot(object = ba4_srt, features = cd_genes)
a_ba4$data
mean(a_ba4$data$pct.exp)


df <- data.frame(modality = rep(c("GPNMB_neg", "GPNMB_pos"),3), 
                 tissue = c("BA4", "BA4", "CB", "CB", "CSC", "CSC"),
                 percentage = c(100- mean(a_ba4$data$pct.exp),
                                mean(a_ba4$data$pct.exp), 
                                100- mean(a_cb$data$pct.exp),
                                mean(a_cb$data$pct.exp), 
                                100- mean(a_csc$data$pct.exp),
                                mean(a_csc$data$pct.exp)))



ggplot(df, aes(x = tissue, y=percentage, fill = modality))+
  geom_bar(stat="identity")+scale_fill_manual(values= mycoloursP[25:50])

```
```{r}
Idents(csc_srt) <- "process_number"
a_csc_s <- DotPlot(object = csc_srt, features = cd_genes)
a_csc_s$data
mean(a_csc_s$data$pct.exp)
a_csc_s$data$Tissue <- "CSC"

Idents(cb_srt) <- "process_number"
a_cb_s <- DotPlot(object = cb_srt, features = cd_genes)
a_cb_s$data
mean(a_cb_s$data$pct.exp)
a_cb_s$data$Tissue <- "CB"

Idents(ba4_srt) <- "process_number"
a_ba4_s <- DotPlot(object = ba4_srt, features = cd_genes)
a_ba4_s$data
mean(a_ba4_s$data$pct.exp)
a_ba4_s$data$Tissue <- "BA4"

b_data <- rbind(a_csc_s$data, 
                a_cb_s$data,
                a_ba4_s$data)

```

```{r}
ggplot(b_data, aes(x = Tissue, y = pct.exp)) +
  geom_boxplot()+ geom_jitter(width = 0.2, aes(colour = id)) +
  theme_bw(19) + ylab("GPNMB+ (%)") + xlab("Tissue")

```

```{r}
lm_rna <- lm(b_data$pct.exp~ b_data$Tissue)
summary(lm_rna)
anova(lm_rna)
```


Conclusion:
I found that automatically counting FMN1 positive cells lead to a large variation, 
and hoped manual counting would alleviate this. But I still observe a lot of
variation among samples. The FMN1 validation is tricky, because it is 
expressed in processes and it is sometimes difficult to decide if a process
belongs to a cell or merely touches it. I think what I can say is that a subset
of oligodendrocytes do express FMN1 and that differences in their percentage 
between snRNAseq and IF may be due to technical differences, sample differences
or inaccuracy in quantifying IF images due to FMN1 being expressed in processes. 
It was particularly difficult to quantify in Spinal cord due to staining of background structures (ECM/ cytoskelleton).

```{r}
sessionInfo()

```