---
title: "combine_val_data_cluster"
author: "Luise A. Seeker"
date: "01/09/2022"
output: html_document
---


# Introduction
With this script I am plotting co-expression of marker genes in the snRNAseq and
validation datasets. 

N.B. This scipts takes variables from my environment.

```{r}
library(gridExtra)

```
Take objects from environment and re-name

1) SPARC OPALIN

```{r}
SPARC_opalin_olig2

#sparc_opalin_long <- long_data


sparc_opalin_seq <- subs_df_snrna_seq

sparc_opalin_seq$marker <- ifelse(sparc_opalin_seq$gene_expr == "SPARC+", "Marker1+",
                                 ifelse(sparc_opalin_seq$gene_expr == "OPALIN+",
                                        "Marker2+", "Double+"))

sparc_opalin_seq$marker <- factor(sparc_opalin_seq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))

sparc_opalin_seq$colabel <- "SPARC OPALIN"


```

2) SPARC RBFOX1

```{r}
sparc_rbfox1_seq <- subs_data

sparc_rbfox1_seq$marker <- ifelse(sparc_rbfox1_seq$gene_expr == "SPARC+", "Marker1+",
                                 ifelse(sparc_rbfox1_seq$gene_expr == "RBFOX1+",
                                        "Marker2+", "Double+"))

sparc_rbfox1_seq$marker <- factor(sparc_rbfox1_seq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))

sparc_rbfox1_seq$colabel <- "SPARC RBFOX1"


# IF data

sp_rb_prot <- sort_red_long

```



3) OPALIN RBFOX1


```{r}



opalin_rbfox1_seq <- subs_df_snrna_seq

opalin_rbfox1_seq$marker <- ifelse(opalin_rbfox1_seq$gene_expr == "OPALIN+", "Marker1+",
                                 ifelse(opalin_rbfox1_seq$gene_expr == "RBFOX1+",
                                        "Marker2+", "Double+"))

opalin_rbfox1_seq$marker <- factor(opalin_rbfox1_seq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))

opalin_rbfox1_seq$colabel <- "OPALIN RBFOX1"

```

4)

```{r}

#FMN1 RBFOX1 data
fmn1_rbfox1


head(subs_subs_data)

# complete long_data inclduding all quantified cells
#fmn1_rbfox1_long <- long_data


#FMN1 RBFOX1 snRNAseq

fmn1_rbfox1_seq <- subs_df_snrna_seq

fmn1_rbfox1_seq$marker <- ifelse(fmn1_rbfox1_seq$gene_expr == "FMN1+", "Marker1+",
                                 ifelse(fmn1_rbfox1_seq$gene_expr == "RBFOX1+",
                                        "Marker2+", "Double+"))

fmn1_rbfox1_seq$marker <- factor(fmn1_rbfox1_seq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


fmn1_rbfox1_seq$colabel <- "FMN1 RBFOX1"

```


5) 
```{r}
# SPARC HCN2 snRNAseq

sp_hc_seq <- subs_data

sp_hc_seq$marker <- ifelse(sp_hc_seq$gene_expr == "SPARC+", "Marker1+",
                                 ifelse(sp_hc_seq$gene_expr == "HCN2+",
                                        "Marker2+", "Double+"))

sp_hc_seq$marker <- factor(sp_hc_seq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


sp_hc_seq$colabel <- "SPARC HCN2"

```

6)
```{r}

# NELL1 PAX3 snRNAseq

ne_pa_seq <- subs_data

ne_pa_seq$marker <- ifelse(ne_pa_seq$gene_expr == "NELL1+", "Marker1+",
                                 ifelse(ne_pa_seq$gene_expr == "PAX3+",
                                        "Marker2+", "Double+"))

ne_pa_seq$marker <- factor(ne_pa_seq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


ne_pa_seq$colabel <- "NELL1 PAX3"


```

```{r}

snRNAseq <- rbind(sparc_opalin_seq, 
                  fmn1_rbfox1_seq, 
                  sparc_rbfox1_seq, 
                  opalin_rbfox1_seq)


snRNAseq <- rbind(snRNAseq[,2:6],
                  sp_hc_seq)

snRNAseq <- rbind(snRNAseq,
                  ne_pa_seq)
```


```{r}
snRNAseq$colabel <- factor(snRNAseq$colabel, 
                           levels = c("SPARC OPALIN",
                                      "SPARC RBFOX1",
                                      "OPALIN RBFOX1",
                                      "FMN1 RBFOX1",
                                      "SPARC HCN2",
                                      "NELL1 PAX3"))

snRNAseq$marker <- factor(snRNAseq$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))

p11 <- ggplot(snRNAseq, aes(x = colabel, y=percentage, fill = marker))+
  geom_bar(stat="identity", position = "fill")+
  scale_fill_manual(values= mycoloursP[24:50])  +
  ylab("Proportion")+
  xlab("") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = -90))

p11
```

```{r}

write.csv(snRNAseq, here("data",
                         "validation_data",
                         "snRNAseq_ol_clust_mark_coexpression.csv"))

```


# combine validation data

```{r}
# prepare sparc opalin olig2 IF data
sp_op_prep <- data.frame(sample_id = sparc_opalin_long$sample_id,
                         donor_id = sparc_opalin_long$donor_id,
                         Tissue = sparc_opalin_long$tissue,
                         keycol = sparc_opalin_long$keycol,
                         valuecol= sparc_opalin_long$valuecol)

sp_op_prep$marker <- ifelse(sp_op_prep$keycol == "Num_SPARC_OLIG2", 
                              "Marker1+",
                              ifelse(sp_op_prep$keycol == "Num_OLIG2_OPALIN", 
                                     "Marker2+",
                                     ifelse(sp_op_prep$keycol == "Num_SPARC_OLIG2_OPALIN",
                                            "Double+", "exclude")))

subs_sd_op <- subset(sp_op_prep, sp_op_prep$marker != "exclude")

subs_sd_op$marker <- as.factor(subs_sd_op$marker)
subs_sd_op$marker <- factor(subs_sd_op$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


subs_sd_op$colabel <- "SPARC OPALIN OLIG2"

# prepare SPARC RNFOX1 OLIF2 IF data


sp_rb_prep <- data.frame(sample_id = sp_rb_prot$sample_id,
                         donor_id = sp_rb_prot$donor_id,
                         Tissue = sp_rb_prot$tissue,
                         keycol = sp_rb_prot$keycol,
                         valuecol= sp_rb_prot$valuecol)

sp_rb_prep$marker <- ifelse(sp_rb_prep$keycol == "Num_SPARC_OLIG2", 
                              "Marker1+",
                              ifelse(sp_rb_prep$keycol == "Num_RBFOX1_OLIG2", 
                                     "Marker2+",
                                     ifelse(sp_rb_prep$keycol == "Num_RBFOX1_SPARC_OLIG2",
                                            "Double+", "exclude")))

subs_sp_rb <- subset(sp_rb_prep, sp_rb_prep$marker != "exclude")

subs_sp_rb$marker <- as.factor(subs_sp_rb$marker)
subs_sp_rb$marker <- factor(subs_sp_rb$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


subs_sp_rb$colabel <- "SPARC RBFOX1 OLIG2"


# prepare OPALIN RBFOX1 OLIG2 IF data

op_rb_prot <- long_data

op_rb_prep <- data.frame(sample_id = op_rb_prot$sample_id,
                         donor_id = op_rb_prot$donor_id,
                         Tissue = op_rb_prot$tissue,
                         keycol = op_rb_prot$keycol,
                         valuecol= op_rb_prot$valuecol)

op_rb_prep$marker <- ifelse(op_rb_prep$keycol == "Num_OPALIN_OLIG2", 
                              "Marker1+",
                              ifelse(op_rb_prep$keycol == "Num_OLIG2_RBFOX1", 
                                     "Marker2+",
                                     ifelse(op_rb_prep$keycol == "Num.OPALIN..OLIG2..RBFOX1",
                                            "Double+", "exclude")))

op_rb_prep <- subset(op_rb_prep, op_rb_prep$marker != "exclude")

op_rb_prep$marker <- factor(op_rb_prep$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


op_rb_prep$colabel <- "OPALIN RBFOX1 OLIG2"

# prepare FMN1 RBFOX1 OLIG2 IF data

fm_rb_prep <- data.frame(sample_id = fmn1_rbfox1_long$sample_id,
                         donor_id = fmn1_rbfox1_long$donor_id,
                         Tissue = fmn1_rbfox1_long$tissue,
                         keycol = fmn1_rbfox1_long$keycol,
                         valuecol= fmn1_rbfox1_long$valuecol)

fm_rb_prep$marker <- ifelse(fmn1_rbfox1_long$keycol == "FMN1_OLIG2", 
                              "Marker1+",
                              ifelse(fm_rb_prep$keycol == "RBFOX1_OLIG2", 
                                     "Marker2+",
                                     ifelse(fm_rb_prep$keycol == "FMN1_RBFOX1_OLIG2",
                                            "Double+", "exclude")))

subs_fm_rb <- subset(fm_rb_prep, fm_rb_prep$marker != "exclude")

subs_fm_rb$marker <- factor(subs_fm_rb$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


subs_fm_rb$colabel <- "FMN1 RBFOX1 OLIG2"

# prepare SPARC HCN2 data 

sp_hc_long <- long_data
sp_hc_long$donor_id <- substr(sp_hc_long$sample_id, 1,8)

sp_hc_prep <- data.frame(sample_id = sp_hc_long$sample_id,
                         donor_id = sp_hc_long$donor_id,
                         Tissue = sp_hc_long$tissue,
                         keycol = sp_hc_long$keycol,
                         valuecol= sp_hc_long$valuecol)

sp_hc_prep$marker <- ifelse(sp_hc_prep$keycol == "Num_OLIG2_SPARC", 
                              "Marker1+",
                              ifelse(sp_hc_prep$keycol == "Num_OLIG2_HCN2", 
                                     "Marker2+",
                                     ifelse(sp_hc_prep$keycol == "Num_OLIG2_SPARC_HCN2",
                                            "Double+", "exclude")))

sp_hc_prep <- subset(sp_hc_prep, sp_hc_prep$marker != "exclude")

sp_hc_prep$marker <- factor(sp_hc_prep$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


sp_hc_prep$colabel <- "SPARC HCN2 OLIG2"


# NELL1 PDGFRA

ne_pd_long <- long_data
ne_pd_long$donor_id <- paste(substr(ne_pd_long$sample_id, 1,8))
ne_pd_long$tissue <- substr(ne_pd_long$sample_id, 10,12)

ne_pd_long <- data.frame(sample_id = ne_pd_long$sample_id,
                         donor_id = ne_pd_long$donor_id,
                         Tissue = ne_pd_long$tissue,
                         keycol = ne_pd_long$keycol,
                         valuecol= ne_pd_long$valuecol)

ne_pd_long$marker <- ifelse(ne_pd_long$keycol == "NELL1_C2", 
                              "Marker1+",
                              ifelse(ne_pd_long$keycol == "PDRFRA_C1", 
                                     "Marker2+",
                                     ifelse(ne_pd_long$keycol == "PDGFRA_NELL1",
                                            "Double+", "exclude")))

ne_pd_long <- subset(ne_pd_long, ne_pd_long$marker != "exclude")

ne_pd_long$marker <- factor(ne_pd_long$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


ne_pd_long$colabel <- "NELL1 PDGFRA"

# PAX3 PDGFRA


pa_pd_long <- long_data
#pa_pd_long$donor_id <- paste(substr(pa_pd_long$sample_id, 1,8))
#pa_pd_long$tissue <- substr(pa_pd_long$sample_id, 10,12)

pa_pdprep <- data.frame(sample_id = pa_pd_long$sample_id,
                         donor_id = pa_pd_long$donor_id,
                         Tissue = pa_pd_long$tissue,
                         keycol = pa_pd_long$keycol,
                         valuecol= pa_pd_long$valuecol)

pa_pdprep$marker <- ifelse(pa_pdprep$keycol == "PAX3_pos", 
                              "Marker1+",
                              ifelse(pa_pdprep$keycol == "PDGFRA_pos", 
                                     "Marker2+",
                                     ifelse(pa_pdprep$keycol == "double_pos",
                                            "Double+", "exclude")))

pa_pdprep <- subset(pa_pdprep, pa_pdprep$marker != "exclude")

pa_pdprep$marker <- factor(pa_pdprep$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


pa_pdprep$colabel <- "PAX3 PDGFRA"


```





# combine datasets

```{r}
#prot_data <- rbind(subs_sd_op, subs_fm_rb, subs_sp_rb, op_rb_prep)
#prot_data <- rbind(prot_data[,2:8],sp_hc_prep)

prot_data <- rbind(prot_data,pa_pdprep)

prot_data$colabel <- factor(prot_data$colabel, 
                           levels = c("SPARC OPALIN OLIG2",
                                      "SPARC RBFOX1 OLIG2",
                                      "OPALIN RBFOX1 OLIG2",
                                      "FMN1 RBFOX1 OLIG2",
                                      "SPARC HCN2 OLIG2",
                                      "NELL1 PDGFRA",
                                      "PAX3 PDGFRA"))

prot_data$marker <- factor(prot_data$marker, levels = c("Marker1+",
                                 "Double+", "Marker2+"))


p1 <- ggplot(prot_data, aes(x = colabel, y=valuecol, fill = marker))+
  geom_bar(stat="identity", position = "fill")+
  scale_fill_manual(values= mycoloursP[24:50])  +
  ylab("Proportion")+
  xlab("") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = -90))

p1


bool_pax3 <- prot_data$colabel == "PAX3 PDGFRA" &
  prot_data$marker == "Marker1+"

subs_prot <- prot_data[!bool_pax3,]

bool_nell1 <- subs_prot$colabel == "NELL1 PDGFRA" &
  subs_prot$marker == "Marker1+"


subs_prot_2 <- subs_prot[!bool_nell1,]


p2 <- ggplot(subs_prot_2, aes(x = colabel, y=valuecol, fill = marker))+
  geom_bar(stat="identity", position = "fill")+
  scale_fill_manual(values= mycoloursP[24:50])  +
  ylab("Proportion")+
  xlab("") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = -90))

p2





```


```{r}
grid.arrange(p11, p1, ncol = 2)

```
```{r}
write.csv(prot_data, here("data",
                         "validation_data",
                         "IF_ol_clust_mark_coexpression.csv"))

```

```{r}

Idents(nad_ol) <- "ol_clusters_named"

oligos <- subset(nad_ol, idents = c("Oligo_A",
                                    "Oligo_B",
                                    "Oligo_C",
                                    "Oligo_D",
                                    "Oligo_E",
                                    "Oligo_F"))

opcs <- subset(nad_ol, idents = c("OPC_A", "OPC_B"))
                 
                 
Idents(opcs) <- "Tissue"
ba4_opcs <-subset(opcs, idents = "BA4")
cb_opcs <-subset(opcs, idents = "CB")
csc_opcs <-subset(opcs, idents = "CSC")


NELL1 <- 'NELL1' 
PDGFRA <- 'PDGFRA' 
NELL1.cutoff <- 1 
PDGFRA.cutoff <- 1 

#BA4
NELL1_cells_ba4 <- length(which(FetchData(ba4_opcs, vars = NELL1) > NELL1.cutoff & 
                                      FetchData(ba4_opcs, vars = PDGFRA) < PDGFRA.cutoff))
PDGFRA_cells_ba4 <- length(which(FetchData(ba4_opcs, vars = PDGFRA) > PDGFRA.cutoff & 
                                      FetchData(ba4_opcs, vars = NELL1) < NELL1.cutoff))
NELL1_PDGFRA_olig2_cells_ba4 <- length(which(FetchData(ba4_opcs, vars = PDGFRA) > PDGFRA.cutoff & 
                                      FetchData(ba4_opcs, vars = NELL1) > NELL1.cutoff))
all_cells_incluster_ba4 <- table(ba4_opcs@active.ident)
NELL1_ba4 <-NELL1_cells_ba4/all_cells_incluster_ba4 * 100 
# Percentage of cells in dataset that express NELL1

PDGFRA_ba4 <- PDGFRA_cells_ba4/all_cells_incluster_ba4 * 100 
#Percentage of cells in dataset that express PDGFRA
double_ba4 <- NELL1_PDGFRA_olig2_cells_ba4/all_cells_incluster_ba4 * 100 
#Percentage of cells in dataset that co-express NELL1 + PDGFRA

#CB
NELL1_cells_cb <- length(which(FetchData(cb_opcs, vars = NELL1) > NELL1.cutoff & 
                                      FetchData(cb_opcs, vars = PDGFRA) < PDGFRA.cutoff))
PDGFRA_cells_cb <- length(which(FetchData(cb_opcs, vars = PDGFRA) > PDGFRA.cutoff& 
                                      FetchData(cb_opcs, vars = NELL1) < NELL1.cutoff))
NELL1_PDGFRA_olig2_cells_cb <- length(which(FetchData(cb_opcs, vars = PDGFRA) > 
                                         PDGFRA.cutoff & FetchData(cb_opcs, vars = NELL1) >
                                         NELL1.cutoff))
all_cells_incluster_cb <- table(cb_opcs@active.ident)
NELL1_cb <- NELL1_cells_cb/all_cells_incluster_cb * 100 # Percentage of cells in dataset that express NELL1
PDGFRA_cb <- PDGFRA_cells_cb/all_cells_incluster_cb * 100 #Percentage of cells in dataset that express PDGFRA
double_cb <- NELL1_PDGFRA_olig2_cells_cb/all_cells_incluster_cb * 100 #Percentage of cells in dataset that co-express NELL1 + PDGFRA

#CSC
NELL1_cells_csc <- length(which(FetchData(csc_opcs, vars = NELL1) > NELL1.cutoff& 
                                      FetchData(csc_opcs, vars = PDGFRA) < PDGFRA.cutoff))
PDGFRA_cells_csc <- length(which(FetchData(csc_opcs, vars = PDGFRA) > PDGFRA.cutoff & 
                                      FetchData(csc_opcs, vars = NELL1) < NELL1.cutoff))
NELL1_PDGFRA_olig2_cells_csc <- length(which(FetchData(csc_opcs, vars = PDGFRA) > 
                                          PDGFRA.cutoff & FetchData(csc_opcs, vars = NELL1) >
                                          NELL1.cutoff))
all_cells_incluster_csc <- table(csc_opcs@active.ident)
NELL1_csc <- NELL1_cells_csc/all_cells_incluster_csc * 100 
# Percentage of cells in dataset that express NELL1
PDGFRA_csc <- PDGFRA_cells_csc/all_cells_incluster_csc * 100 
#Percentage of cells in dataset that express PDGFRA
double_csc <- NELL1_PDGFRA_olig2_cells_csc/all_cells_incluster_csc * 100 
#Percentage of cells in dataset that co-express NELL1 + PDGFRA

df_snrna_seq <- data.frame(Tissue = c(rep("BA4", 4), 
                                     rep("CB", 4),
                                     rep("CSC", 4)),
                           gene_expr = rep(c("NELL1+", 
                                             "PDGFRA+",
                                             "NELL1+PDGFRA+",
                                             "Other OPCs"), 3),
                           percentage = c(as.numeric(NELL1_ba4), 
                                          as.numeric(PDGFRA_ba4),
                                          as.numeric(double_ba4),
                                          100 - (as.numeric(NELL1_ba4)+ 
                                          as.numeric(PDGFRA_ba4)+
                                          as.numeric(double_ba4)),
                                          
                                          as.numeric(NELL1_cb), 
                                          as.numeric(PDGFRA_cb),
                                          as.numeric(double_cb),
                                          100 - (as.numeric(NELL1_cb)+ 
                                          as.numeric(PDGFRA_cb)+
                                          as.numeric(double_cb)),
                                          
                                          
                                          as.numeric(NELL1_csc), 
                                          as.numeric(PDGFRA_csc),
                                          as.numeric(double_csc),
                                          100 - (as.numeric(NELL1_csc)+ 
                                          as.numeric(PDGFRA_csc)+
                                          as.numeric(double_csc))))




ggplot(df_snrna_seq, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity")+scale_fill_manual(values= mycoloursP[24:50])


subs_data <- subset(df_snrna_seq, df_snrna_seq$gene_expr != "Other OPCs")

ggplot(subs_data, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity")+scale_fill_manual(values= mycoloursP[24:50])

ggplot(subs_data, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity", position = "fill")+scale_fill_manual(values= mycoloursP[24:50])
```


```{r}

nell1_data <- subs_data

nell1_data$coexpr <- "NELL1 PDGFRA"

nell1_data$marker <- ifelse(nell1_data$gene_expr == "NELL1+", "Marker1+",
                            ifelse(nell1_data$gene_expr == "PDGFRA+", "PDGFRA+",
                                   "Double+"))

```

PAX3 OPCs

```{r}
PAX3 <- 'PAX3' 
PDGFRA <- 'PDGFRA' 
PAX3.cutoff <- 1 
PDGFRA.cutoff <- 1 

#BA4
PAX3_cells_ba4 <- length(which(FetchData(ba4_opcs, vars = PAX3) > PAX3.cutoff & 
                                      FetchData(ba4_opcs, vars = PDGFRA) < PDGFRA.cutoff))
PDGFRA_cells_ba4 <- length(which(FetchData(ba4_opcs, vars = PDGFRA) > PDGFRA.cutoff & 
                                      FetchData(ba4_opcs, vars = PAX3) < PAX3.cutoff))
PAX3_PDGFRA_olig2_cells_ba4 <- length(which(FetchData(ba4_opcs, vars = PDGFRA) > PDGFRA.cutoff & 
                                      FetchData(ba4_opcs, vars = PAX3) > PAX3.cutoff))
all_cells_incluster_ba4 <- table(ba4_opcs@active.ident)
PAX3_ba4 <-PAX3_cells_ba4/all_cells_incluster_ba4 * 100 
# Percentage of cells in dataset that express PAX3

PDGFRA_ba4 <- PDGFRA_cells_ba4/all_cells_incluster_ba4 * 100 
#Percentage of cells in dataset that express PDGFRA
double_ba4 <- PAX3_PDGFRA_olig2_cells_ba4/all_cells_incluster_ba4 * 100 
#Percentage of cells in dataset that co-express PAX3 + PDGFRA

#CB
PAX3_cells_cb <- length(which(FetchData(cb_opcs, vars = PAX3) > PAX3.cutoff & 
                                      FetchData(cb_opcs, vars = PDGFRA) < PDGFRA.cutoff))
PDGFRA_cells_cb <- length(which(FetchData(cb_opcs, vars = PDGFRA) > PDGFRA.cutoff& 
                                      FetchData(cb_opcs, vars = PAX3) < PAX3.cutoff))
PAX3_PDGFRA_olig2_cells_cb <- length(which(FetchData(cb_opcs, vars = PDGFRA) > 
                                         PDGFRA.cutoff & FetchData(cb_opcs, vars = PAX3) >
                                         PAX3.cutoff))
all_cells_incluster_cb <- table(cb_opcs@active.ident)
PAX3_cb <- PAX3_cells_cb/all_cells_incluster_cb * 100 # Percentage of cells in dataset that express PAX3
PDGFRA_cb <- PDGFRA_cells_cb/all_cells_incluster_cb * 100 #Percentage of cells in dataset that express PDGFRA
double_cb <- PAX3_PDGFRA_olig2_cells_cb/all_cells_incluster_cb * 100 #Percentage of cells in dataset that co-express PAX3 + PDGFRA

#CSC
PAX3_cells_csc <- length(which(FetchData(csc_opcs, vars = PAX3) > PAX3.cutoff& 
                                      FetchData(csc_opcs, vars = PDGFRA) < PDGFRA.cutoff))
PDGFRA_cells_csc <- length(which(FetchData(csc_opcs, vars = PDGFRA) > PDGFRA.cutoff & 
                                      FetchData(csc_opcs, vars = PAX3) < PAX3.cutoff))
PAX3_PDGFRA_olig2_cells_csc <- length(which(FetchData(csc_opcs, vars = PDGFRA) > 
                                          PDGFRA.cutoff & FetchData(csc_opcs, vars = PAX3) >
                                          PAX3.cutoff))
all_cells_incluster_csc <- table(csc_opcs@active.ident)
PAX3_csc <- PAX3_cells_csc/all_cells_incluster_csc * 100 
# Percentage of cells in dataset that express PAX3
PDGFRA_csc <- PDGFRA_cells_csc/all_cells_incluster_csc * 100 
#Percentage of cells in dataset that express PDGFRA
double_csc <- PAX3_PDGFRA_olig2_cells_csc/all_cells_incluster_csc * 100 
#Percentage of cells in dataset that co-express PAX3 + PDGFRA

df_snrna_seq <- data.frame(Tissue = c(rep("BA4", 4), 
                                     rep("CB", 4),
                                     rep("CSC", 4)),
                           gene_expr = rep(c("PAX3+", 
                                             "PDGFRA+",
                                             "PAX3+PDGFRA+",
                                             "Other OPCs"), 3),
                           percentage = c(as.numeric(PAX3_ba4), 
                                          as.numeric(PDGFRA_ba4),
                                          as.numeric(double_ba4),
                                          100 - (as.numeric(PAX3_ba4)+ 
                                          as.numeric(PDGFRA_ba4)+
                                          as.numeric(double_ba4)),
                                          
                                          as.numeric(PAX3_cb), 
                                          as.numeric(PDGFRA_cb),
                                          as.numeric(double_cb),
                                          100 - (as.numeric(PAX3_cb)+ 
                                          as.numeric(PDGFRA_cb)+
                                          as.numeric(double_cb)),
                                          
                                          
                                          as.numeric(PAX3_csc), 
                                          as.numeric(PDGFRA_csc),
                                          as.numeric(double_csc),
                                          100 - (as.numeric(PAX3_csc)+ 
                                          as.numeric(PDGFRA_csc)+
                                          as.numeric(double_csc))))




ggplot(df_snrna_seq, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity")+scale_fill_manual(values= mycoloursP[24:50])


subs_data <- subset(df_snrna_seq, df_snrna_seq$gene_expr != "Other OPCs")

ggplot(subs_data, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity")+scale_fill_manual(values= mycoloursP[24:50])

ggplot(subs_data, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity", position = "fill")+scale_fill_manual(values= mycoloursP[24:50])
```


```{r}

PAX3_data <- subs_data

PAX3_data$coexpr <- "PAX3 PDGFRA"

PAX3_data$marker <- ifelse(PAX3_data$gene_expr == "PAX3+", "Marker1+",
                            ifelse(PAX3_data$gene_expr == "PDGFRA+", "PDGFRA+",
                                   "Double+"))

comb_data <- rbind(nell1_data, PAX3_data)


subs_dat_opc <- subset(comb_data, comb_data$marker != "Marker1+")
```


```{r}
ggplot(subs_dat_opc, aes(x = coexpr, y=percentage, fill = marker))+
  geom_bar(stat="identity", position = "fill")+
  scale_fill_manual(values= mycoloursP[27:50]) + 
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = -90))

```

```{r}
ggplot(subs_data, aes(x = Tissue, y=percentage, fill = gene_expr))+
  geom_bar(stat="identity", position = "fill")+scale_fill_manual(values= mycoloursP[24:50])
```

OPC validation data

Taken from environment (68_Validation_Nell1_PDGFRA.Rmd and
58_Validation_PDGFRA_PAX3)
```{r}
n_dat <- data.frame(donor_id = nell1_subs_long$donor_id, 
                    tissue = nell1_subs_long$tissue,
                    sample_id = nell1_subs_long$sample_id,
                    keycol = nell1_subs_long$keycol,
                    valuecol= nell1_subs_long$valuecol,
                    staining = nell1_subs_long$staining)

n_dat$marker <- ifelse(n_dat$keycol == "PDGFRA_NELL1",
                       "Double+", "PDGFRA+")

p_dat <- data.frame(donor_id = pax3_subs_long$donor_id, 
                    tissue = pax3_subs_long$tissue,
                    sample_id = pax3_subs_long$sample_id,
                    keycol = pax3_subs_long$keycol,
                    valuecol= pax3_subs_long$valuecol,
                    staining = pax3_subs_long$staining)

p_dat$marker <- ifelse(p_dat$keycol == "double_pos",
                       "Double+", "PDGFRA+")


nell_pax <- rbind(n_dat, p_dat)

ggplot(nell_pax, aes(x = staining , 
                          y = valuecol, 
                          fill = marker)) +
  geom_bar(position="fill", stat="identity")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values= mycoloursP[16:30])+
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = -90))

```

```{r}

sessionInfo()
```